Effects of knockdown of autophagy pathway genes on C. elegans longevity are highly condition dependent

Version 1

Abstract

Autophagy is proposed to protect against aging by clearing damaged cellular constituents. In line with this several life-extending interventions in model organisms show some degree of autophagy dependence. In C. elegans, inhibiting autophagy can shorten, lengthen or have no effect on lifespan. Differences between published findings likely reflect variability in experimental conditions. Here we investigate the condition dependence of effects on lifespan of RNA-mediated interference (RNAi) knockdown of autophagy pathway components. Effects on several interventions causing a strong Age (increased lifespan) phenotype were examined, including mutation of daf-2 (insulin/IGF-1 receptor). Factors varied included daf-2 mutant allele class, atg gene, temperature and presence of 5-fluoro-2’-deoxyuridine (FUDR). Effects on lifespan of atg RNAi proved to be highly condition dependent. Notably, for most atg genes tested lifespan was not usually reduced more in the long-lived mutant than in the wild-type control. Greater suppression was seen at 20°C for certain atg genes with daf-2(e1368) but not daf-2(e1370). At 25°C, little reduction in lifespan was seen. However, atg-18 knockdown behaved differently, suppressing daf-2 Age under all conditions, suggesting possible pleiotropic action. FUDR at a high concentration caused knockdown of several atg genes to increase lifespan. Thus, depending on experimental conditions, atg knockdown can increase, decrease or have no effect on daf-2 Age. The lack of suppression of Age by atg RNAi under most conditions questions the importance of autophagy in daf-2 Age. Moreover, condition dependence of effects creates a risk of possible condition selection bias.

Introduction

A long-standing theory is that senescence (aging) is largely a consequence of the accumulation of random molecular damage caused by, among other things, reactive oxygen species (Beckman and Ames, 1998; Harman, 1956; Murphy, 2023; Shore and Ruvkun, 2013; Szilard, 1959). This view predicts that mechanisms of somatic maintenance, particularly those that prevent accumulation of damaged cellular constituents, will decelerate the aging process. One somatic maintenance function viewed as a potential longevity-assurance process is autophagy (specifically macroautophagy), which effects lysosome-dependent degradation of cellular constituents, including damaged matter (Aman et al., 2021).

The possible anti-aging role of autophagy has been extensively tested in the short-lived nematode Caenorhabditis elegans, and supporting evidence found with respect to several life-extending interventions, including reduced insulin/IGF-1 signaling (IIS), reduced germline signaling, and dietary restriction (Hansen et al., 2018). These studies were principally performed in the 2000s; however, during the same period, falsification tests of the molecular damage theory, particularly that relating to oxidative damage, led to some uncertainty about its validity (Gems and Doonan, 2009; Perez et al., 2009; Shields et al., 2021). Meanwhile, an alternative theoretical framework emerged, based on the evolutionary theory of aging (Arnold and Rose, 2023; Williams, 1957), arguing that senescence is largely the consequence of genetically-determined, programmatic mechanisms (Blagosklonny, 2006; de Magalhães and Church, 2005; Gems, 2022; Maklakov and Chapman, 2019), and very much so in C. elegans (Gems and de la Guardia, 2013; Gems et al., 2021; Pires da Silva et al., 2024).

One form of programmatic aging involves costly programs: genetically-determined processes that degrade somatic tissues as a by-product of wider, fitness-promoting processes (Gems and Kern, 2024; Gems et al., 2021). In some cases this involves biomass repurposing, where biomass of one tissue is broken down by autophagic processes to release molecular constituents to support functions in another. This occurs to a high degree in semelparous organisms in the process of reproductive death (suicidal reproductive effort), as in many monocarpic plants, and semelparous fish such as Pacific salmon (Gems et al., 2021).

Several lines of evidence support the hypothesis that reproductive death occurs in C. elegans hermaphrodites (Gems et al., 2021; Kern et al., 2023). This includes a putative costly program in which intestinal biomass is broken down and repurposed to support production of yolk, that is then vented to support larval growth, leading to intestinal atrophy (a senescent pathology) (Ezcurra et al., 2018; Kern et al., 2021; Sornda et al., 2019). Notably, such biomass repurposing is supported by autophagy, as evidenced by deceleration of intestinal atrophy and yolk pool formation when autophagy is inhibited (Benedetto and Gems, 2019; Ezcurra et al., 2018). Thus, in this particular context autophagy appears to be promoting rather than preventing senescence.

These developments warrant a careful re-examination of the evidence that autophagy is protective against aging in C. elegans. In this study we re-examine the question of whether the longevity of daf-2 insulin/IGF-1 receptor mutants is autophagy dependent. Here the principal form of past evidence involves epistasis tests, where effects on lifespan of reduction of function of genes encoding proteins effecting autophagy is compared in the wild type (N2) and daf-2 mutants. Findings from 8 previous studies involving 46 epistasis experiments are summarised in Table S1. Although it is widely believed that autophagy is essential for daf-2 mutant longevity (Meléndez et al., 2003), scrutiny of the results of past tests raises doubts about the strength of this claim.

Careful consideration of these prior studies identifies six distinct limitations, as follows. (1) A life-shortening effect of inhibition of autophagy does not necessarily indicate its role in the normal aging process, or in daf-2 mutant longevity. (2) If autophagy is inhibited during development as well as adulthood, a life-shortening effect could result from disruption of normal development. (3) The claim that daf-2 longevity is autophagy dependent requires evidence that the life-shortening effect in daf-2 is significantly greater than in wild type, e.g. using Cox proportional hazard analysis, and this is rarely performed. (4) If effects of reducing autophagy are condition dependent, this introduces a potential bias: a risk that investigators might unwittingly select conditions where the data generated supports a role of autophagy in longevity - what may be referred to as condition selection bias. Potential determinative conditions that have varied across studies include choice of autophagy-determining gene to inhibit, of daf-2 mutant allele, and ambient temperature. (5) 5-fluoro-2’-deoxyuridine (FUDR) is sometimes used to facilitate C. elegans lifespan assays by preventing progeny hatching; this could potentially affect test outcomes, as shown in other contexts (Aitlhadj and Sturzenbaum, 2010; Anderson et al., 2016; Burnaevskiy et al., 2018; Van Raamsdonk and Hekimi, 2011; Zhao et al., 2019). (6) A final, straightforward issue is whether a given finding proves to be reproducible in subsequent reports under, seemingly, the same conditions.

Of 46 prior tests (Table S1) only 4 present clear evidence that reducing autophagy shortens lifespan more in daf-2 than in the wild type. Regarding one of these four instances, the effect of bec-1 RNAi on daf-2(e1370) at 15°C (Meléndez et al., 2003), a subsequent study did not replicate this finding (Hars et al., 2007). In two other cases the weaker daf-2(mu150) allele was used (Hansen et al., 2008; Patel et al., 2008). In two further studies where large reductions in daf-2 lifespan were observed (Chang et al., 2017; Minnerly et al., 2017) knockdown of atg-18 was used. In a 2009 study where effects of knockdown of 14 different autophagy genes was tested, adult-limited atg-18 RNAi was one of only 2/14 that significantly reduced lifespan in the wild type (Hashimoto et al., 2009), suggesting possible atg-18 idiosyncrasy. The 2009 study includes more than half of all of the prior tests (28/46); strikingly, in only 1/14 genes (atg-4.1) did adult-limited RNAi significantly reduce daf-2(e1370) lifespan, while for 3/14 genes (atg-9, bec-1 and unc-51) it increased daf-2 lifespan (Hashimoto et al., 2009). One possible reason for the lack of observed life-shortening effects given autophagy knockdown is its use of FUDR at a high concentration.

In this study we assess the condition-dependence of the effects of RNAi knockdown of autophagy genes on C. elegans longevity. To this end we have tested the effects of RNAi of six genes in the autophagy pathway on longevity in two different daf-2 mutants, and a glp-1 germlineless mutant, at two temperatures. We have also assessed effects of FUDR and prevention of bacterial infection. Our results provide a robust foundation of evidence relating to possible autophagy dependence of daf-2 and glp-1 longevity. They suggest a weak and highly-condition dependent contribution of autophagy to daf-2 and glp-1 Age. This implies that prevention of damage accumulation by autophagy is, at most, a minor determinant of daf-2 longevity.

Results

Effects of atg RNAi on daf-2 Age vary with daf-2 allele

We first tested whether effects on lifespan of inhibiting autophagy depend upon daf-2 allele severity. Two daf-2 mutants were examined. daf-2(e1368) is a class 1 (less pleiotropic) mutant where adults show normal behavior at both 20°C and 25°C, while daf-2(e1370) is a class 2 (more pleiotropic) mutant where at 25°C adults become paralyzed and cease feeding (Gems et al., 1998). Knockdown of autophagy was performed by RNAi from the L4 (late larval) stage, to exclude possible confounding, life-shortening disruption of normal development. RNAi was performed for six genes at several stages of the autophagy process: initiation (atg-13), membrane nucleation (atg-9, bec-1), phagophore formation (atg-2, atg-18) and elongation (atg-4.1) (Figure 1A). 5/6 genes were those examined in our previous study (Ezcurra et al., 2018), with bec-1 added because it was the subject of several earlier studies (Table S1).

Effects of atg RNAi on longevity are highly variable and condition dependent.

(A) The autophagy pathway, and genes tested in this study. (B-D) Effects at 20°C. (B) Effects on N2 (wild type). (C) Effects on daf-2(e1368). (D) Effects on daf-2(e1370). (F, G) Effects at 25°C. (F) Effects on N2. (G) Effects on daf-2(e1368). (H) Effects of bec-1 RNAi on daf-2(e1370) at 15°C. (B-G summed data, N = 2; for individual trials and statistical comparisons, see Table S2, S5.

A methodological note: for tests of effects of a given intervention on C. elegans lifespan an often-applied standard is to include 3 biological replicates. This is true of several recent studies where the effect of knockdown of a single atg gene on daf-2 longevity was studied (Minnerly et al., 2017; Wilhelm et al., 2017; Yang et al., 2024). However, given that the present condition dependence study effectively performs this test in 18 different ways, involving RNAi of 6 atg genes, 2 daf-2 mutants and 2 temperatures, N = 2 biological replicates were judged to be sufficient to draw robust conclusions; similarly, an earlier study of RNAi 14 atg genes under two conditions used 2-3 biological replicates (Hashimoto et al., 2009); for an overview of N sizes in previous studies, see Table S1.

Trials were performed at 20°C, and used gfp RNAi as a negative control. In wild-type C. elegans (N2), statistically significant reductions were sometimes seen: atg-9, atg-13 and, particularly, atg-18 RNAi shortened lifespan in both trials (summed data, N = 2, means: −13.7%, p < 0.0001, −12.4%, p = 0.0002, −31%, p < 0.0001, respectively; log rank test) (Figure 1B, Table S2; for all raw mortality data, see Supplemental Dataset 1). For survival curves comparing RNAi effects of individual atg genes on the three genotypes, see Figure 2.

Degree of suppression of daf-2 longevity by atg gene RNAi differs greatly between daf-2 alleles and atg RNAi treatments (20°C), cf. Figure 1B-D (same data).

Summed data, N = 2; for individual trials and statistical comparisons, see Table S2. All trials were performed in parallel, i.e. all lifespans are directly comparable.

RNAi of all six atg genes consistently reduced lifespan in the daf-2(e1368) mutant, with effects ranging from −10.5% (p = 0.0009) for atg-4.1 to −50.4% (p < 0.0001) for atg-18 (Figure 1C, Table S2). To assess whether atg RNAi reduced lifespan more in daf-2 mutants than in the wild type, Cox proportional hazard (CPH) analysis was used. This showed a significantly greater effect in daf-2(e1368) for 4/6 genes (exceptions: atg-4.1, bec-1) (Table S2). These findings are in line with the earlier observation that bec-1 and vps-34 RNAi shortened the lifespan of the daf-2(mu150) class 1 mutant but not of N2 at 20°C (Hansen et al., 2008).

To test whether atg RNAi is able to fully suppress daf-2(e1368) Age, the lifespans of N2 and daf-2 subjected to atg RNAi were compared. Under all six atg RNAi conditions, daf-2(e1368) still significantly increased mean lifespan, from +15.7% (p = 0.0031) for atg-13 to +60.9% (p < 0.0001) for atg-4.1 (summed data, 20°C, Table S2). This could imply either that daf-2(e1368) Age is not fully autophagy dependent, or that it is but autophagy is not fully suppressed by the RNAi treatment.

In daf-2(e1370), atg RNAi had far more modest effects on lifespan than in daf-2(e1368). Summed data showed significant reductions in lifespan resulting from RNAi of atg-9, atg-13 and atg-18 only, with the latter again causing the largest reduction: −8.3% (p = 0.012), −17.8% (p = 0.0034), and −29.2% (p < 0.0001), respectively (Figure 1D, Table S2). However, effects were in no instance significantly greater than in N2 (CPH analysis, Table S2), thus failing to provide evidence that e1370 longevity is mediated by autophagy. Taken together, these results could imply that at 20°C autophagy contributes to longevity in weaker but not stronger daf-2 mutants, possibly in class 1 but not class 2 mutants.

RNAi effects varied between atg genes, with atg-2 and atg-4.1 RNAi effects weaker, and atg-18 RNAi effects generally stronger than the rest. One possible cause of this variability is differences between the plasmid vectors used for RNAi in terms of how efficiently they destroy their target mRNA. To investigate this possibility, RNAi of each atg gene was performed in N2 animals, and transcript levels measured by RT–qPCR and analyzed using the ΔΔCt method, normalized to a gfp RNAi control, with fold changes calculated as 2^−ΔΔCt (n = 4 independent trials). Effects of RNAi on mRNA level varied greatly, from no detected reduction in atg-2 to an 86% reduction in atg-18 (Fig. S1A; Table S3; Supplementary dataset 2). Pairwise comparisons of ΔΔCt values showed that, atg-2 aside, there were no significant differences between RNAi treatment effects on mRNA levels, apart from a greater effect of atg-18 RNAi when compared to either atg-4.1 and atg-9 RNAi (Fig. S1B).

Comparing effects of RNAi on mRNA levels and lifespan, the only clear correspondence between the former and the latter involved atg-2 and atg-4.1 RNAi, which in N2 had no significant effect on either (Figure 1B, Fig. S1A; Table S2). However for the remaining 4 genes, there was no clear correspondence, though atg-18 mRNA levels appear to be the lowest, in line with its greater effect on lifespan (Figure 1B-D, Fig. S1A; Table S2, Table S3). These results imply that RNAi efficacy and perhaps also gene-specific issues contribute to variation in atg RNAi effects on lifespan (discussed further below). While mRNA levels after RNAi under the various other conditions tested were not assayed, reduced IIS (including daf-2(e1370)) intensifies the RNAi response (Wang and Ruvkun, 2004), thus lack of effect on lifespan in daf-2(e1370) is very unlikely to reflect suppression of mRNA knockdown.

Effect of atg RNAi on daf-2 Age is temperature dependent

Results of a previous study performed at 25°C appear to show no greater reduction in lifespan in daf-2(e1370) compared to daf-2(+) after autophagy knockdown, even when using the atg-18(gk378) deletion mutation (Toth et al., 2008) (Table S1), suggesting possible temperature dependence of atg RNAi effects. To explore this further, in parallel to tests at 20°C, we also compared effects of atg RNAi on lifespan in N2 and daf-2(e1368) at 25°C. Effects on daf-2(e1370) were not tested, partly because this mutant ceases feeding at 25°C (Gems et al., 1998) which would be expected to interfere with RNAi by feeding.

At 25°C, atg RNAi did not shorten N2 lifespan for any of the six genes tested, not even atg-18 (summed data; Figure 1E, Table S2), i.e. culture at 25°C suppressed the life-shortening effect of atg RNAi in N2. Also notable is that the increases in lifespan with atg-2 and atg-13 RNAi at 25°C, described in our earlier study (Ezcurra et al., 2018), were not reproduced (discussed below).

At 25°C only atg-18 RNAi significantly reduced daf-2(e1368) mean lifespan, by 15.5% (p < 0.0001) (summed data, Figure 1F, Table S2), a reduction that was significantly greater than in N2 (p = 0.0012, CPH, summed data only; Table S2). For survival curves comparing RNAi effects of individual atg genes on the two genotypes, see Fig. S2. In one case, atg-9, RNAi slightly increased daf-2 lifespan (+9.8%, p = 0.011, summed data).

The initial tests showing that bec-1 RNAi reduces daf-2(e1370) lifespan were performed at 15°C (Meléndez et al., 2003); moreover, a greater N2 life-shortening effect of bec-1 RNAi at 15°C than 20°C has been reported (Chen et al., 2019). Taken together with the weaker RNAi effects at 25°C observed here, this suggested that stronger effects might be seen at 15°C. To explore this we compared effects of bec-1 RNAi from L4 on N2 and daf-2(e1370) at 15°C and 20°C (N = 2). However, no suppression of daf-2(e1370) Age by bec-1 RNAi was seen at either temperature (Figure 1G; Table S5), consistent with findings of an earlier study performed at 15°C (Hars et al., 2007) (Table S1).

FUDR can alter the effect of atg RNAi on lifespan

Since the 1980s FUDR, a thymidylate synthase inhibitor and anti-cancer drug, has sometimes been added to C. elegans survival trials to prevent progeny production (Gandhi et al., 1980; Mitchell et al., 1979). Notably, two previous reports that observed increases in C. elegans lifespan given atg RNAi employed FUDR. Our own study saw increases in N2 lifespan after atg-2 and atg-13 RNAi from L4 with 15 µM FUDR (Ezcurra et al., 2018). Another study that saw increases in N2 lifespan given adult-limited atg-7, atg-9, bec-1 and unc-51 RNAi used FUDR at a higher concentration, 800 μM (Hashimoto et al., 2009) (E. Nishida, personal communication).

To test for FUDR-dependent effects, we compared the impact of atg gene RNAi (atg-2, atg-4.1, atg-9, atg-13, atg-18 and bec-1) on N2 lifespan at 20°C with 0, 15 or 800 µM FUDR (N = 2). With 0 or 15 µM FUDR, only shortening of lifespan was seen, and FUDR did not significantly alter the effects of RNAi (CPH, Figure 3A,B, Table S6).

FUDR but not infection alters outcome on atg gene RNAi (20°C).

(A-C) Effects of 0 μM, 15 μM and 800 µM FUDR. (D) No alteration by kanamycin of atg-13 RNAi effect on N2 lifespan. Summed data, N = 2; for individual trials and statistical comparisons, see Table S6, S7.

Addition of 800 µM FUDR increased the lifespan of the gfp RNAi negative control by 61.7% (p < 0.0001, Table S6), perhaps due to prevention of bacterial proliferation, which can otherwise shorten C. elegans lifespan (Garigan et al., 2002; Gems and Riddle, 2000). In the presence of 800 µM FUDR, atg-9 and atg-13 RNAi increased lifespan, by +12.5% (p = 0.014) and +8.7% (p = 0.031), respectively (summed data; Figure 3C, Table S6). Moreover, the life-shortening effect of bec-1 RNAi, seen with 0 or 15 µM FUDR, was absent. Again, atg-18 RNAi robustly reduced lifespan under all three conditions (Figure 3C, Table S6). The results using 800 µM FUDR are broadly in line with those of Hashimoto et al. (2009), where atg-9 and atg-13 RNAi increased lifespan and atg-18 was one of only 2/14 genes tested where adult-limited RNAi decreased N2 lifespan. This suggests that the increases in lifespan after atg RNAi reported in that study could have reflected its use of 800 µM FUDR (Hashimoto et al., 2009).

atg RNAi shortens lifespan in the absence of bacterial infection

Under standard laboratory culture conditions, C. elegans lifespan is limited by infection by the E. coli food source, such that prevention of bacterial proliferation substantially increases lifespan (Garigan et al., 2002; Gems and Riddle, 2000). The preceding results could imply that 800 µM FUDR suppresses life-shortening effects of atg RNAi by preventing bacterial infection. Xenophagy is generally protective against infection in C. elegans (Balla et al., 2019; Jia et al., 2009). Thus, reduction in lifespan given atg gene knockdown could reflect increased susceptibility to infection.

To probe this hypothesis, we compared effects on N2 lifespan of atg-13 RNAi at 20°C in the absence or presence of the antibiotic kanamycin (Kan), to suppress bacterial infection. In the absence of Kan, atg-13 RNAi caused a slight reduction in lifespan in these trials that did not reach statistical significance (Figure 3D, Table S7), in contrast to other trials in this study (Table S2, S6). Application of Kan extended C. elegans lifespan (+27.1%, p < 0.0001, summed data, Table S7), as previously observed (Garigan et al., 2002), and in its presence atg-13 RNAi resulted in the same slight reduction in lifespan (Figure 3D, Table S7). These results suggest that life-shortening effects of atg RNAi are not solely attributable to increased susceptibility to E. coli infection. Moreover, they do not indicate marked condition dependency in atg RNAi effects on lifespan with respect to E. coli proliferative status.

atg-18 RNAi robustly suppresses glp-1(e2141) Age

We next explored more widely the reproducibility and condition dependence of the requirement for autophagy of C. elegans longevity. Prevention of germline development by laser surgery or mutation increases lifespan in C. elegans hermaphrodites (Arantes-Oliveira et al., 2002; Hsin and Kenyon, 1999; Pires da Silva et al., 2024). Prior tests for possible autophagy dependence of such longevity have largely used the temperature-sensitive glp-1(e2141) germline proliferation mutant, which is fertile at 15°C but sterile and with greatly reduced germline development at 25°C. A key study reported strong and reproducible suppression of glp-1 Age by RNAi of five autophagy-related genes, including atg-18 and bec-1 (Lapierre et al., 2011) (previous findings summarized in Table S8).

We first tested the effect of atg RNAi on glp-1 longevity with animals raised from L1 until L4 stage at 25°C, and maintained at 20°C thereafter, similar to previous studies (Table S8). Again, effects of atg-2, atg-4.1, atg-9, atg-13, atg-18 and bec-1 RNA were tested. In the RNAi control glp-1 increased mean lifespan by +50.9% (p < 0.0001, summed data, Table S9). glp-1 lifespan was significantly reduced by atg-2, atg-4.1, atg-9, atg-18 and bec-1 RNAi (but not atg-13 RNAi), and effects were greater in glp-1 than N2 in all cases except atg-4.1 (Figure 4A, B, Figure 5, Table S9). However, suppression was only robust (of a large magnitude) for atg-2 and atg-18 RNAi (Figure 4B, Figure 5). Under all six atg RNAi conditions, glp-1 still significantly increased mean lifespan, from +4.12% (p = 0.0003) for atg-18 to +43.8% (p < 0.0001) for atg-4.1 (summed data, 20°C, Table S9). This could imply either that glp-1 Age is not fully autophagy dependent, or that it is but autophagy is not fully suppressed by the RNAi treatment.

Degree of suppression of glp-1(e2141) longevity by atg gene RNAi differs greatly between genes.

(A, B) 20°C. (A) Effects on N2. (B) Effects on glp-1. (C, D) 25°C. (C) Effects on N2. (D) Effects on glp-1. Summed data, N = 2-5; for individual trials and statistical comparisons, see Table S9. (E) Overview of effects of suppression of Age by RNAi of genes specifying autophagy. Dark blue, reduction of daf-2 or glp-1 Age that is significantly greater than in N2. Light blue, reduction of daf-2 or glp-1 Age that is not significantly greater than in N2. R, robust suppression, i.e. strong but incomplete suppression of longevity, defined here as <30% increase in mean lifespan in daf-2 or glp-1 relative to N2, with RNAi, and as evident from Figure 2, Figure 5, and Fig. S2.

Degree of suppression of glp-1(e2141) longevity by atg gene RNAi differs greatly between genes (20°C), cf. Figure 4A,B (same data).

Summed data, N = 2-5; for individual trials and statistical comparisons, see Table S9.

In tests with daf-2(e1368) life-shortening effects of atg RNAi were largely absent at 25°C (apart from atg-18) (Figure 1F, Table S2). We therefore wondered whether temperature might also influence the outcome of atg RNAi treatment in glp-1 mutants. To test this we examined RNAi effects on lifespan at 25°C. Under these conditions, glp-1 lifespan was significantly reduced by only atg-2, atg-4.1, and atg-18 RNAi (Figure 4C, D, Fig. S3, Table S9), and effects were significantly greater in glp-1 than N2 only with atg-2 and atg-18 RNAi (CPH analysis, Table S9). In summary, of the six genes tested only atg-2 and atg-18 RNAi robustly suppressed glp-1 Age, and RNAi effects were weaker at 25°C.

For an overview of the effects of RNAi on daf-2 and glp-1, see Figure 4E. In 11 of the 30 conditions tested RNAi knockdown of autophagy caused a greater reduction in lifespan in the long-lived mutant. In 7/30 the RNAi effect was robust, i.e. the mutant longevity was largely suppressed. This suppression was highly condition dependent, differing according to gene knocked down, temperature, daf-2 allele used, and between daf-2 and glp-1. Notably, only atg-18 RNAi robustly suppressed longevity in both daf-2 and glp-1 mutants.

Why are effects of atg-18 RNAi stronger than those of other atg genes?

The particularly strong life-shortening effects of atg-18 RNAi could reflect either a greater reduction in autophagy, or the presence in addition to its effects on autophagy of pleiotropic effects not directly related to autophagy. Given that longevity due to either daf-2 mutation or germline loss are wholly dependent on the FOXO transcription factor DAF-16 (Hsin and Kenyon, 1999; Kenyon et al., 1993), we wondered whether atg-18 RNAi might inhibit DAF-16. To probe this two approaches were taken. First, we used a constitutive dauer formation assay. High population density and food depletion causes C. elegans larvae to form developmentally arrested dauer larvae (Cassada and Russell, 1975). daf-2 mutants undergo constitutive dauer arrest (the Daf-c phenotype) in a temperature-sensitive manner, and this is fully suppressed by daf-16(−) (Riddle et al., 1981). However, using a sensitive assay (daf-2(m41), 22°C, giving a mix of dauer and non dauers), we detected no reduction in the number of dauers formed given atg gene RNAi (atg-2, atg-13 and atg-18) (Fig. S4A).

Second, we tested a GFP reporter for a gene whose expression is elevated in daf-2 mutants in a daf-16-dependent manner (ftn-1), using strains previously constructed (Ackerman and Gems, 2012). GFP levels were compared in daf-2(m577) and daf-16(mgDf50); daf-2 backgrounds. As expected, GFP levels were higher in daf-2 than in daf-16; daf-2 (Fig. S4B). RNAi did not suppress the daf-2-induced increase of ftn-1::gfp expression (Fig. S4B). These findings argue against a pleiotropic effect of atg-18 on DAF-16 function.

Inhibiting autophagy does not reduce vitellogenin accumulation

Finally, we further investigated the hypothesis that autophagy promotes biomass repurposing in C. elegans. Inhibition of yolk synthesis or of autophagy delays intestinal atrophy and yolk pool accumulation, suggesting that intestinal biomass is repurposed for yolk synthesis (Benedetto and Gems, 2019; Ezcurra et al., 2018; Sornda et al., 2019). In principle, this could involve repurposing into yolk protein (vitellogenin) or yolk lipid. To test the former possibility, wild-type hermaphrodites or fog-2(q71) (feminization of germline) mutant females were subjected to atg RNAi. The fog-2 mutant, which lacks self-sperm and so lays no eggs, was used to avoid possible effects of atg RNAi on fertility, reduction of which can increase vitellogenin levels within nematodes (Sornda et al., 2019).

For none of the six atg genes tested did RNAi detectably reduce yolk protein levels, either in N2 hermaphrodites or fog-2 females (Figure 6). This implies that autophagic machinery, including that in the intestine, does not enhance yolk protein production. Thus, if intestinal biomass repurposing occurs, then it likely supports yolk secretion or yolk lipid production.

Absence of effect of atg RNAi on vitellogenin accumulation.

Fold change of yolk proteins YP170, YP115 and YP88 in N2 and fog-2 normalized to gfp day 1 (N = 3). In no case is vitellogenin level significantly different to that in the gfp RNAi control at any time point (two-way ANOVA, Table S10). For raw data for vitellogenin levels see Supplementary Dataset 3.

Discussion

Overall, the effects of reducing atg gene function described here are ambiguous with respect to the role of autophagy in daf-2 or glp-1 longevity, neither strongly supporting or excluding it. However, they are consistent with a role of autophagy in the longevity of weaker, class 1 daf-2 alleles at lower temperatures (Figure 1C,D,F,G, Table S2). Class 1 allele-limited suppression of daf-2 Age has been seen previously, for example in epistasis tests with daf-12 (encoding a dafachronic acid receptor) (Gems et al., 1998; Larsen et al., 1995) and skn-1 (Nrf2-like transcription factor) (Tullet et al., 2008). This could reflect a role of autophagy in longevity assurance limited to conditions of mild IIS reduction, or sensitivity to differential effects of daf-2 on distinct downstream signaling outputs, such as phosphatidylinositol 3-kinase and Ras signaling (Patel et al., 2008).

atg RNAi suppresses daf-2(e1368) Age at 20°C but not 25°C (Figure 1C,F, Table S2). Given that for hypomorphic daf-2 alleles (such as e1368 and e1370) many mutant traits, including Age, show some degree of temperature sensitivity (Gems et al., 1998; Riddle and Albert, 1997), the absence of suppression at 25°C may reflect the increased severity of the daf-2(e1368) mutant phenotype at this higher temperature. However, life-shortening effects of atg RNAi on N2 are also reduced at 25°C, suggesting that additional mechanisms may also mediate such temperature sensitivity.

Resolving discrepancies between past studies of autophagy dependence of daf-2 Age

The increasing non-reproducibility of many experimental findings, that has been referred to as the reproducibility crisis, is a problem that particularly afflicts biological and biomedical research (Baker, 2016; Lithgow et al., 2017; Prinz et al., 2011; Ritchie, 2020; Voelkl et al., 2020). A strength of C. elegans as an experimental model is the relative ease with which such discrepancies can be resolved, at least in principle. This is thanks to the use of standardized culture conditions across the C. elegans research community, and of nematode strains based on the same isogenic wild-type strain (N2), plus the relatively low cost and short duration of experiments.

Regarding lifespan assays in particular, possible reasons for discrepant findings include clearly identifiable differences in experimental design, such as use or not of FUDR. Less obvious causes include cryptic variation in genetic background (Zhao et al., 2019), or subtle differences in culture conditions (e.g. due to batch variation in Bacto Peptone, a constituent of nematode growth medium) (Petrascheck, 2014).

Our findings potentially resolve several discrepancies between earlier studies relating to the possible role of autophagy in daf-2 mutant longevity. Previous studies found that inhibiting autophagy either suppressed or enhanced daf-2 Age, or had no effect (Table S1). One study showing suppression used the weak class 1 allele daf-2(mu150) (Hansen et al., 2008). This is consistent with our observation that, atg-18 aside, atg RNAi can suppress Age in a class 1 but not a class 2 allele (Figure 1C,D Table S2).

Does bec-1 RNAi suppress daf-2(e1370) Age? The initial test suggesting this was performed at 15°C, and employed RNAi by injection of the mothers of assayed individuals (Meléndez et al., 2003). In a subsequent study at 15°C, in which both mothers and adult progeny were exposed to bec-1 RNAi by feeding, daf-2(e1370) lifespan appeared not to be shortened (Hars et al., 2007). Several further trials under different conditions did not observe a life-shortening effect either (Hashimoto et al., 2009; Toth et al., 2008). In the present study, we saw no shortening of daf-2(e1370) lifespan in summed data with adult-limited bec-1 RNAi by feeding, at either 15°C or 20°C (Figure 1D,G, Fig. S3, Table S2, S5). Taken together with earlier evidence, our observations suggest that suppression of daf-2(e1370) Age by bec-1 RNAi is not a readily reproducible finding.

Several studies reported lifespan extension following atg RNAi (Ezcurra et al., 2018; Hashimoto et al., 2009). Here we were able to reproduce this effect for several atg genes by applying high dose FUDR (800 µM), thus recapitulating findings by Hashimoto et al (2009), and potentially accounting for the life span increases seen in that study; it was only subsequent to that study that evidence emerged of the capacity of FUDR to alter effects of interventions that impact lifespan (Aitlhadj and Sturzenbaum, 2010; Anderson et al., 2016; Burnaevskiy et al., 2018; Van Raamsdonk and Hekimi, 2011; Zhao et al., 2019).

Regarding issues with replicating published findings, in an earlier study we observed increases in N2 lifespan given atg-2 and atg-13 RNAi (15 µM FUDR present) (Ezcurra et al., 2018). However, this finding proved not to be robust to replication (Figure 3B, Table S6), perhaps reflecting inherent effect variability, condition dependence or a combination of the two. This again illustrates the value, as with the effects of bec-1 RNAi on daf-2(e1370) Age, of repeated verification of effects of interventions on C. elegans lifespan.

As with daf-2, our tests with the glp-1 germline proliferation mutant did not clearly support the view that mutant longevity is autophagy dependent. However, consistent with an earlier study (Lapierre et al., 2011), we observed that atg-18 and bec-1 RNAi reduce glp-1 lifespan more than in wild-type at 20°C.

Why does atg-18 RNAi more strongly suppress daf-2 Age?

Of the six autophagy-determining genes tested here, RNAi of atg-18 showed a greater capacity to suppress both daf-2 and glp-1 Age than the other five. atg-18 appears to have replaced bec-1 as the gene of choice for autophagy-related epistasis studies in C. elegans (Chang et al., 2017; Minnerly et al., 2017). Notably, atg-18 was one of only 2/14 autophagy genes tested where RNAi during development caused a high level (>50%) of larval growth arrest or lethality (Hashimoto et al., 2009). The more marked effects of atg-18 RNAi could reflect either greater inhibition of autophagy, or pleiotropy in which processes other than autophagy are altered.

Regarding pleiotropy, several proteins in the canonical autophagy pathway have recently been found to participate in other processes. For example, ATG8 (LGG-1 in C. elegans) functions in trafficking of single-membrane organelles (Nieto-Torres et al., 2021), and C. elegans ATG-16.2 contributes to neuronal exopher formation via its WD40 domain (Yang et al., 2024). In an as yet unexplained instance of atg gene idiosyncrasy, mutation of atg-16.2, atg-18 or bec-1 retards the cell cycle in C. elegans germline cells, while that of atg-7 does not (Ames et al., 2017).

Concerning possible atg-18 pleiotropy, several studies suggest this possibility. ATG-18 is a predicted WIPI (WD repeat protein interacting with phosphoinositides) family member. In humans there are four WIPI proteins, WIPI1 - WIPI4; C. elegans atg-18 is more closely related to WIPI1/WIPI2, while epg-6 more closely resembles WIPI3/WIPI4 (Lu et al., 2011). Notably, deletion mutations of atg-18 and epg-6 appear to cause similar reductions in levels of autophagy, yet only atg-18 shortens lifespan (Takacs et al., 2019).

Next, a C. elegans studies of rescue of longevity and fat storage in daf-2(e1370); atg-18(gk378) double mutants by tissue-specific expression of atg-18(+) revealed a major role of atg-18 in food-sensing chemosensory neurons, potentially reflecting a vesicle trafficking role in neurosecretion (possibly of neuropeptides) (Jia et al., 2019; Minnerly et al., 2017). Whether such a role is unique to atg-18, or shared by other atg genes remains unclear.

Third, recent evidence suggests that ATG-18 activates HLH-30/TFEB (transcription factor EB). HLH-30 is a master transcriptional regulator of autophagy and lysosomal biogenesis, that becomes nuclear localized in daf-2(e1370) and glp-1(e2141) mutants, and whose inhibition reduces longevity in both contexts (Lapierre et al., 2013; Lin et al., 2018; Wong et al., 2023). Notably, among genes up-regulated upon over-expression of atg-18, ones with HLH-30 elements in their promoters are over-represented, and similarly in genes down-regulated in an atg-18(gk378) mutant. Also, heat shock-induced nuclear localization of HLH-30 is suppressed by atg-18 RNAi, as is longevity resulting from atg-18 over-expression (Schmauck-Medina et al., 2026).

Inhibition of HLH-30 by atg-18 RNAi could potentially explain its unusually strong life-shortening effects. If correct, this could reflect greater suppression of autophagy, but also, in principle, other effects of HLH-30 inhibition. HLH-30 affects a variety of traits, including adult reproductive diapause (Gerisch et al., 2020), proteostasis (Shalash et al., 2025), sex differences in immunity (Sohn et al., 2025), and resistance to various insults, including infectious pathogens (El-Houjeiri et al., 2019; Visvikis et al., 2014; Wani et al., 2021), starvation (Harvald et al., 2017; O’Rourke and Ruvkun, 2013; Settembre et al., 2013), and oxidative and heat stress (Lin et al., 2018).

While in at least some of these cases autophagy likely contributes to hlh-30-mediated effects, given that this transcription factor also promotes expression of many genes unrelated to autophagy (Chen et al., 2017; Lin et al., 2018; Shalash et al., 2025; Sohn et al., 2025; Visvikis et al., 2014), other functions may also be involved. For example, in response to infection HLH-30 can activate expression of signaling (including IIS), autophagy-related and immunity-related genes, and both of latter can contribute to infection resistance (Chen et al., 2017; Sohn et al., 2025; Visvikis et al., 2014). In conclusion, if the more severe effects of atg-18 RNAi on lifespan reported here are attributable to reduced HLH-30 activity, this could well reflect greater reduction of autophagy. However, it remains possible that other functions controlled by HLH-30 also play a role, e.g. relating to immunity (Chen et al., 2017; Visvikis et al., 2014) and protein quality control (Shalash et al., 2025). Also notable is that HLH-30 regulates gene expression combinatorially with DAF-16 (Lin et al., 2018), mutation of which fully suppresses daf-2(rf) longevity (Kenyon et al., 1993).

Autophagy in biomass repurposing during programmatic aging

Autophagic processes play a major role in tissue degeneration related to biomass repurposing in semelparous organisms (that reproduce once and then die), particularly plants (Avila-Ospina et al., 2014; Gems et al., 2021). Previous studies support the hypothesis that intestinal biomass is repurposed for synthesis of yolk that, subsequent to egg laying, is vented to support larval development (Kern et al., 2021), and that autophagy facilitates this biomass conversion (Ezcurra et al., 2018; Sornda et al., 2019). If late-life mortality is promoted by intestinal atrophy, then preventing it should extend lifespan. Consistent with this, intestinal atrophy and yolk production are suppressed in daf-2 mutants (Depina et al., 2011; Ezcurra et al., 2018), and blocking yolk production both retards intestinal atrophy and modestly increases lifespan (Ezcurra et al., 2018; Murphy et al., 2003; Sornda et al., 2019). This could imply that inhibiting autophagy, by retarding intestinal atrophy could extend lifespan under some conditions, and several earlier studies observed life extension after atg RNAi (Ezcurra et al., 2018; Hashimoto et al., 2009; Wilhelm et al., 2017).

The results of the present study are in line with the view that intestinal atrophy, though a salient feature of C. elegans senescence, by itself contributes only weakly to late-life mortality. Consistent with this, inhibition of daf-2 using auxin-induced degradation of DAF-2 protein can strongly increasing lifespan even when initiated only at advanced ages, long after intestinal atrophy has occurred, and without any detectable reversal of major senescent pathologies (Molière et al., 2024; Venz et al., 2021). One possibility is that atg RNAi has antagonistic effects on late-life mortality: modestly reducing it due to suppression of intestinal atrophy (as seen when vitellogenin synthesis is inhibited) (Ezcurra et al., 2018; Murphy et al., 2003; Sornda et al., 2019), but also increasing it, due to disruption of essential cellular functions.

The gut-to-yolk biomass repurposing hypothesis drew particularly on the observation that blocking yolk protein (vitellogenin) synthesis or autophagy reduced intestinal atrophy and pseudocoelomic yolk pool size (Ezcurra et al., 2018; Sornda et al., 2019); given that yolk pools contain yolk protein and lipid (Ezcurra et al., 2018; Garigan et al., 2002; Herndon et al., 2002; Palikaras et al., 2017), such putative biomass repurposing could increase production of yolk protein, yolk lipid or both. Here we tested whether atg RNAi reduces yolk protein levels, and found that it did not (Figure 6). If the hypothesis is correct, then this would support the view that promotion by autophagy of repurposing of intestinal biomass into yolk involves yolk lipid rather than yolk protein, or overall yolk secretion. In principle, this could involve mass export of stored lipid, either by means of lipophagy, or perhaps secretory autophagy (Ponpuak et al., 2015). As an aside, we note that there is evidence that vitellogenin levels can affect autophagy (Seah et al., 2016).

Concluding remarks

This study reassesses the claim that daf-2 Age is autophagy dependent. While the results do not entirely prove or disprove this claim, they clearly show that evidence of dependence varies greatly with context, and suggest that autophagy only contributes substantially to daf-2 Age given a weak reduction in IIS.

One limitation in establishing the role of autophagy in daf-2 mutant longevity is the difficulty of reliably measuring its level of activity. While several assays using fluorescent reporters to identify putative autophagosomes suggest possible increases in autophagy levels (Chang et al., 2017; Hansen et al., 2008; Meléndez et al., 2003), daf-2 mutants also show marked reductions in both protein synthesis rate (Depuydt et al., 2013) and turnover rate (Dhondt et al., 2016; Visscher et al., 2016), including reduced ATG-18 turnover rate (Visscher et al., 2016). This is more consistent with a reduction rather than an increase in autophagy, and in line with observed suppression of intestinal atrophy in daf-2 mutants (Ezcurra et al., 2018).

One issue raised by condition dependence of a given effect is that it creates a risk of condition selection bias, where choice of experimental conditions can bias the outcome of an experiment. Where condition dependence has been identified, condition selection bias may be avoided by selection of multiple conditions to test a given hypothesis. In the present case, this could entail comparing effects on daf-2(e1368) and daf-2(e1370) at 20°C, ideally without FUDR use, and RNAi tests with multiple atg genes, at least until the reason for the idiosyncratic effects of atg-18 RNAi are understood. General questions relating to how to deal with condition dependency have been discussed previously (Munafò and Davey Smith, 2018; Voelkl et al., 2020).

A second, wider issue is that condition dependence risks generating a situation where different research articles make conflicting claims, supporting different views held by different groups of researchers. For science to progress effectively it is necessary for research communities to resolve discrepancies between published findings, though the work required can be tedious, a task akin to washing the dishes in a communal household. It requires identifying and flagging findings that are either condition dependent or, seemingly, unreproducible and, where possible, distinguishing the two. For C. elegans at least, such a tidying process is highly feasible, as is use of methodologies devised to reduce such discrepancies (Driscoll et al., 2025; Lucanic et al., 2017; Petrascheck and Miller, 2017), which is a particular virtue of this model organism.

Materials and Methods

Culture methods and strains

C. elegans maintenance was performed using standard protocols (Brenner, 1974). Unless otherwise stated, all strains were on nematode growth media (NGM, containing Bacto Peptone) with plates seeded with E. coli OP50 to provide a food source. An N2 hermaphrodite stock recently obtained from the Caenorhabditis Genetics Center was used as wild type (N2H) (Zhao et al., 2019). Genotypes of most mutants used are as described in Wormbase (www.wormbase.org). Strains used included CB4027 glp-1(e2141), GA633 daf-2(m577); wuIs177 [Pftn-1::gfp lin-15(+)], GA643 daf-16(mgDf50); daf-2(m577); wuIs177 [Pftn-1::gfp lin-15(+)]], GA1930 daf-2(e1370), GA1945 daf-2(m41), GA1960 daf-2(e1368), and JK574 fog-2(q71) V.

Constitutive dauer larva formation assay

daf-2(m41) animals were maintained on RNAi feeding strains for at least two generations prior to analysis. For the dauer larva formation assay, performed at 22°C, 10 L4-stage hermaphrodites were transferred onto RNAi plates and allowed to lay eggs for approximately 6 h, after which adults were removed. Dauers were scored at 166 h.

Epifluorescence microscopy

Nematodes were anaesthetized with 10 μl 2 mM levamisole on 2% agar pads prior to imaging. For imaging, we used either a Zeiss Axioskop 2 plus microscope with Hamamatsu ORCA-ER digital camera C4742-95 and Volocity 6.3 software (Macintosh version) for image acquisition; or an ApoTome.2 Zeiss microscope with a Hamamatsu digital camera C13440 ORCA-Flash4.0 V3 and Zen software for image acquisition.

RNA-mediated interference (RNAi)

RNAi by feeding was performed using RNAi plasmids transformed into E. coli OP50(xu363) as previously described (Xiao et al., 2015). Bacterial transformants were selected on LB agar plates containing 10 µg/ml tetracycline and 25 µg/ml carbenicillin. Inserts of all RNAi feeding clones were confirmed by sequencing. Origins of plasmids in RNAi feeding strains: atg-13, bec-1: Ahringer library (Kamath et al., 2003); atg-2, atg-4.1, atg-9: Vidal library (Rual et al., 2004). As an atg-18 RNAi feeding clone was not available in local RNAi libraries, a custom atg-18 RNAi plasmid was generated by cloning the target sequence into the L4440 vector using primers gcctccacttcctgttgaag and gagactcttttcgtcggca.

For RNAi induction, bacteria were grown for 16 h at 37°C with shaking (200 rpm) in 5 ml LB supplemented with 25 µg/ml carbenicillin. Overnight cultures were diluted 1:100 into fresh LB containing 25 µg/ml carbenicillin and grown for a further 4 h or until reaching an OD₆₀₀ of ∼0.4. Expression of double-stranded RNA was induced by addition of 1 mM IPTG, and cultures were incubated for an additional 1 h. Following induction, cultures were allowed to cool to room temperature, concentrated five-fold by centrifugation, and seeded onto NGM plates supplemented with 25 µg/ml carbenicillin and 1 mM IPTG. Seeded plates were allowed to dry at room temperature before use.

To minimize IPTG degradation, RNAi plates were stored in foil-lined boxes. Typically, unseeded plates were stored at 4°C for up to one month, while seeded plates were used within two weeks. RNAi treatment was initiated from the L4 larval stage unless otherwise stated.

Survival analysis

Nematodes were maintained at a density of 25-30 per plate, and transferred daily during the egg laying period, and every 6-7 days thereafter. The L4 stage was defined as day 0. Mortality was scored every 1-2 days, with worms scored as alive if they showed any movement, either spontaneously or in response to gentle touch with a worm pick.

RNA extraction, cDNA synthesis and RT-qPCR

Approximately 100–200 day 4 adult animals per treatment group were lysed in 250 µl of TRIzol™ reagent by vortexing for 10 min at 4°C, followed by a 10 min incubation on ice; this cycle was repeated three times. RNA was purified using the RNeasy Mini Kit (QIAGEN, Cat. No. 74104) according to the manufacturer’s instructions, including on-column DNase digestion using the RNase-Free DNase Set (QIAGEN, Cat. No. 79256). cDNA was synthesized from 100 ng of total RNA using the SuperScript™ First-Strand Synthesis System (Thermo Fisher Scientific, Cat. No. 10684803).

PCR primers were designed using Primer-BLAST with the following parameters: maximum primer length of 20 nucleotides, PCR product size of 70–200 bp, and exon–exon spanning where possible. Primer efficiency and R² values were determined from standard curves to ensure suitability for quantitative analysis and comparable amplification efficiencies between primer pairs. Primers were used at a final concentration of 2 µM.

Quantitative PCR was performed using Fast SYBR™ Green chemistry on a QuantStudio™ 6 Flex Real-Time PCR System (Applied Biosystems) with 384-well plates and a total reaction volume of 10 µl. Cycling conditions were 95°C for 20 s, followed by 40 cycles of 95°C for 1 s and 65°C for 20 s, with fluorescence data acquisition during the annealing/extension step. Melt curve analysis was performed following amplification to confirm product specificity. A full list of primers used in this study is provided in Table S11. mRNA levels were normalized to mRNA levels from two housekeeping genes, cdc-42 and pmp-3 (Hoogewijs et al., 2008).

Electrophoresis of C. elegans yolk protein

N2 and fog-2(q71) worms were synchronized by performing an egg lay and allowing nematodes to grow at 20°C until they reached the L4 stage. L4 worms were then transferred to fresh seeded RNAi plates and maintained under standard culture conditions at 20°C. Five worms were harvested at day 1, day 4 and day 7 into microcentrifuge Eppendorf tubes containing 10 μl M9 buffer. Before running the gel, 10 μl of 2x Laemmli sample buffer was added. Samples were incubated at 70°C and vortexed periodically for 15 min. Samples were then incubated at 95°C and vortexed periodically for 5 min, and were centrifuged at 6,000 rpm for 15 min.

10 μl of sample was loaded into wells of Invitrogen NuPAGE Bis-Tris protein gels. 5 μl of CozyHi pre-stained protein ladder was loaded at the left side of the gels. The running buffer used was 5% XT MOPS (Bio-Rad). The gels were then run at 150V for 2 hr. Gels were removed from cassette and placed in 100 ml of Coomassie staining solution overnight. Gels were then washed with distilled water three times, then placed into destaining solution and soaked for 40 min. Gels were then washed with distilled water and stored at 4°C until imaged. Gels were imaged and saved as 8 bit grayscale TIF files. Images of gels were analysed using Fiji software. Bands of interest (e.g. myosin, YP170, YP115, YP88) were selected manually based on their molecular weight, and intensity of the bands was measured and exported to Microsoft Excel for further analysis.

Data analysis and statistics

Data were plotted using ePrism 9.0 (GraphPad Software, Boston, MA, USA) or Jupyter Notebook using Python with the matplotlib, pandas and NumPy libraries. Statistical tests were performed on raw data using either Prism 9 or JMP Pro 15 (JMP Statistical Discovery LLC, Cary, NC, USA) unless otherwise stated. The specific tests and post hoc corrections performed are described in the figure legends. To detect alterations in lifespan, the log rank test was used. To compare the magnitude of changes in lifespan, Cox Proportional Hazard (CPH) analysis was used. RT-qPCR data was analyzed using the ΔΔCt method and t-tests. To compare yolk protein levels, a two-way ANOVA was used. For lifespan trials, no statistical methods were used to predetermine sample size. The experiments were not randomized. The investigators were not blinded to allocation during experiments and outcome assessment.

eLife Assessment

This study presents a valuable finding on the condition dependence of autophagy-mediated lifespan regulation in C. elegans. The evidence is solid, as the data broadly support the main claims, although variability between biological replicates and limited mechanistic exploration leave some conclusions less firmly established. The work will be of interest to researchers studying autophagy, ageing, and intracellular trafficking.

Reviewer #1 (Public review):

Summary:

Hsiung et al. investigated whether the effects of autophagy gene knockdown on the lifespan of long-lived C. elegans mutants depend on experimental conditions. The authors first compiled published data on autophagy-dependent lifespan regulation in daf-2 and wild-type backgrounds, highlighting that prior results are notably inconsistent and likely context-dependent. They then systematically tested the lifespan effects of RNAi knockdown of six autophagy genes (atg-2, atg-4.1, atg-9, atg-13, atg-18, and bec-1) in wild-type (N2), daf-2 (reduced insulin/IGF-1 signalling), and glp-1 (germlineless) animals, while varying temperature, daf-2 allele, FUDR concentration, and bacterial infection status.

The key findings are as follows. In wild-type animals, lifespan suppression by most autophagy gene knockdowns was more pronounced at 20{degree sign}C than at 25{degree sign}C, where little or no effect was observed. In daf-2 mutants, stronger lifespan suppression was seen in the weaker daf-2(e1368) allele at 20{degree sign}C, but not in the stronger daf-2(e1370) allele, and effects were largely absent at 25{degree sign}C. In glp-1 mutants, four of six gene knockdowns suppressed lifespan to a greater extent than in N2, though again in a temperature-dependent manner. FUDR at a high concentration (800 µM) abolished the life-shortening effects of most knockdowns and, in the case of atg-9 and atg-13, led to lifespan extension. Kanamycin treatment to eliminate bacterial proliferation did not fully account for the lifespan effects, suggesting that increased susceptibility to infection is not the primary mechanism. The authors also tested the programmed aging hypothesis that autophagy promotes lifespan reduction through biomass repurposing, but found no changes in vitellogenin levels upon knockdown of any of the six genes.

Altogether, among all genes tested, atg-18 knockdown produced the strongest and most consistent lifespan suppression across nearly all conditions, including both daf-2 and glp-1 backgrounds. The authors probed whether atg-18 acts through the FOXO transcription factor DAF-16 by examining dauer formation and ftn-1 expression, but found no evidence for this, suggesting a DAF-16-independent mechanism.

Strengths:

The primary strength of this work lies in its systematic and comprehensive approach to dissecting how experimental variables influence the outcome of autophagy-lifespan epistasis tests. The compilation of prior data alongside the authors' own multi-condition dataset is a genuinely useful resource for the field. The study raises a timely and important point about condition selection bias, which is relevant not only to autophagy research but to C. elegans aging studies more broadly. The finding that atg-18 behaves distinctly from other autophagy genes across all conditions is noteworthy and opens avenues for future mechanistic work.

Weaknesses:

Despite its breadth, the study has several weaknesses that limit the strength of some conclusions.

(1) Variability in control lifespan data. The N2 lifespan values under ostensibly identical conditions (e.g., GFP RNAi at 20{degree sign}C) differ substantially across experiments (compare Tables S2, S5, S6, S7, and S9). Since N2 serves as the baseline for calculating whether the effect is greater in long-lived mutants via Cox proportional hazard (CPH) analysis, this variability in controls directly affects the reliability of those comparisons.

(2) Limited biological replication. Most experiments were performed with only two biological replicates. In several cases, the two replicates yield contradictory outcomes: one showing significant lifespan suppression and the other showing no effect or even extension. The authors combine these into cumulative datasets for analysis, which, while not incorrect in principle, may obscure genuine irreproducibility. Given that the central message of the paper concerns variability and condition dependence, additional replication would have substantially strengthened confidence in the reported results.

(3) Low sample sizes in individual trials. A number of lifespan assays were conducted with only 40-50 worms per replicate, and in some cases, as few as 30. Such sample sizes are below the standard commonly used in the C. elegans aging field and are likely to contribute to the variability observed.

(4) RNAi efficacy measured only in N2 at 20{degree sign}C. The authors demonstrated that atg-2 and atg-4.1 RNAi did not significantly reduce target mRNA levels, which may explain their weaker lifespan effects. However, these same RNAi treatments significantly affected lifespan in several other conditions (e.g., daf-2(e1368) at 20{degree sign}C, glp-1 at 20{degree sign}C and 25{degree sign}C, and N2 with 15 µM FUDR). Measuring RNAi efficacy across different genetic backgrounds and conditions would be needed to properly interpret these variable results.

(5) Incomplete mechanistic exploration. The investigation of why atg-18 knockdown has uniquely strong effects was limited to DAF-16. Given published evidence that atg-18 may regulate HLH-30/TFEB, a master transcriptional regulator of autophagy and lysosomal biogenesis, testing whether atg-18 specifically affects HLH-30 nuclear localisation or activity could have provided valuable mechanistic insight and would distinguish atg-18 from the other genes tested.

Reviewer #2 (Public review):

Summary:

This study examines how genes involved in cellular recycling (autophagy) influence lifespan under different experimental conditions. The findings help clarify why previous studies have reported conflicting results about whether blocking autophagy shortens or extends lifespan. The work will be of interest to researchers studying aging and cellular stress responses, particularly those using model organisms.

Strengths:

The findings are valuable, as they help resolve inconsistencies within a specific subfield of aging research. The evidence presented is solid, as the data broadly support the primary claims of the study. In addition, the discussion is thorough and thoughtfully integrates the findings within the broader context of the field.

Weaknesses:

Additional functional validation would further strengthen the conclusions.

Author response:

The following is the authors’ response to the original reviews.

Public Reviews:

Reviewer #1 (Public review):

Summary:

Hsiung et al. investigated whether the effects of autophagy gene knockdown on the lifespan of long-lived C. elegans mutants depend on experimental conditions. The authors first compiled published data on autophagy-dependent lifespan regulation in daf-2 and wild-type backgrounds, highlighting that prior results are notably inconsistent and likely context-dependent. They then systematically tested the lifespan effects of RNAi knockdown of six autophagy genes (atg-2, atg4.1, atg-9, atg-13, atg-18, and bec-1) in wild-type (N2), daf-2 (reduced insulin/IGF-1 signalling), and glp-1 (germlineless) animals, while varying temperature, daf-2 allele, FUDR concentration, and bacterial infection status.

The key findings are as follows. In wild-type animals, lifespan suppression by most autophagy gene knockdowns was more pronounced at 20°C than at 25°C, where little or no effect was observed. In daf-2 mutants, stronger lifespan suppression was seen in the weaker daf-2(e1368) allele at 20°C, but not in the stronger daf-2(e1370) allele, and effects were largely absent at 25°C. In glp-1 mutants, four of six gene knockdowns suppressed lifespan to a greater extent than in N2, though again in a temperature-dependent manner. FUDR at a high concentration (800 µM) abolished the life-shortening effects of most knockdowns and, in the case of atg-9 and atg-13, led to lifespan extension. Kanamycin treatment to eliminate bacterial proliferation did not fully account for the lifespan effects, suggesting that increased susceptibility to infection is not the primary mechanism. The authors also tested the programmed aging hypothesis that autophagy promotes lifespan reduction through biomass repurposing, but found no changes in vitellogenin levels upon knockdown of any of the six genes.

Altogether, among all genes tested, atg-18 knockdown produced the strongest and most consistent lifespan suppression across nearly all conditions, including both daf-2 and glp-1 backgrounds. The authors probed whether atg-18 acts through the FOXO transcription factor DAF-16 by examining dauer formation and ftn-1 expression, but found no evidence for this, suggesting a DAF-16-independent mechanism.

We thank the reviewer for their meticulous review of the manuscript. Responding to their comments (particularly those not in the public review) has improved it substantially.

Strengths:

The primary strength of this work lies in its systematic and comprehensive approach to dissecting how experimental variables influence the outcome of autophagy-lifespan epistasis tests. The compilation of prior data alongside the authors' own multi-condition dataset is a genuinely useful resource for the field. The study raises a timely and important point about condition selection bias, which is relevant not only to autophagy research but to C. elegans aging studies more broadly. The finding that atg-18 behaves distinctly from other autophagy genes across all conditions is noteworthy and opens avenues for future mechanistic work.

Weaknesses:

Despite its breadth, the study has several weaknesses that limit the strength of some conclusions.

(1) Variability in control lifespan data. The N2 lifespan values under ostensibly identical conditions (e.g., GFP RNAi at 20°C) differ substantially across experiments (compare Tables S2, S5, S6, S7, and S9). Since N2 serves as the baseline for calculating whether the effect is greater in long-lived mutants via Cox proportional hazard (CPH) analysis, this variability in controls directly affects the reliability of those comparisons.

Such inter-trial variability in N2 lifespan is not unusual in studies of C. elegans aging, ostensibly identical conditions notwithstanding, and its causes are unknown. A careful C. elegans lifespan study from 2017 compared results of tests performed independently at 3 sites under similar conditions. This showed that inter-trial variation occurred mainly at each site over time, rather than between sites (Lucanic et al., 2017). We note that the data for our study was gathered over a 5-year period. One possibility is that this long time frame may have contributed to the variability seen.

Regarding statistical tests (whether the log-rank test to assess differences between pairs of populations, or the CPH test to assess differences between effects of treatments under two conditions): comparisons were made either of data within individual trials, or for pooled data. Thus, valid comparisons were made. We did not compare controls from one trial with treatments from another, which would have yielded misleading results.

(2) Limited biological replication. Most experiments were performed with only two biological replicates. In several cases, the two replicates yield contradictory outcomes: one showing significant lifespan suppression and the other showing no effect or even extension. The authors combine these into cumulative datasets for analysis, which, while not incorrect in principle, may obscure genuine irreproducibility. Given that the central message of the paper concerns variability and condition dependence, additional replication would have substantially strengthened confidence in the reported results.

This is a natural issue to raise. In tests of an effect of a given treatment on C. elegans lifespan, a minimum of 3 trials is standard, and in this lab as a rule we follow this convention. However, after careful consideration during the design stages we opted not to do so for this particular study. Our reasons are set out in the manuscript as follows.

“A methodological note: for tests of effects of a given intervention on C. elegans lifespan an often-applied standard is to include 3 biological replicates. This is true of several recent studies where the effect of knockdown of a single atg gene on daf-2 longevity was studied (Minnerly et al., 2017; Wilhelm et al., 2017; Yang et al., 2024). However, given that the present condition dependence study effectively performs this test in 18 different ways, involving RNAi of 6 atg genes, 2 daf-2 mutants and 2 temperatures, N = 2 biological replicates were judged to be sufficient to draw robust conclusions; similarly, an earlier study of RNAi 14 atg genes under two conditions used 2-3 biological replicates (Hashimoto et al., 2009); for an overview of N sizes in previous studies, see Table S1.”

While this approach has yielded robust broad conclusions (e.g. that atg gene RNAi generally does not suppress daf-2(e1370) Age), it is true that for any one given treatment (say, effects of atg-9 RNAi on daf-2(e1370) Age) one may not draw conclusions with a high degree of confidence, and we do not do so. We therefore argue that in a study of this nature, as for instance in a whole genome RNAi screen for lifespan effects, it is reasonable and expedient to drop below the 3 replicates minimum standard; here we agree with the Nishida lab’s similar judgement, and from that study too robust conclusions may be drawn.

(3) Low sample sizes in individual trials. A number of lifespan assays were conducted with only 40-50 worms per replicate, and in some cases, as few as 30. Such sample sizes are below the standard commonly used in the C. elegans aging field and are likely to contribute to the variability observed.

Please see our response to point 2, which in essence responds to this concern.

(4) RNAi efficacy measured only in N2 at 20°C. The authors demonstrated that atg-2 and atg-4.1 RNAi did not significantly reduce target mRNA levels, which may explain their weaker lifespan effects. However, these same RNAi treatments significantly affected lifespan in several other conditions (e.g., daf-2(e1368) at 20°C, glp-1 at 20°C and 25°C, and N2 with 15 µM FUDR). Measuring RNAi efficacy across different genetic backgrounds and conditions would be needed to properly interpret these variable results.

The study would indeed be strengthened by inclusion of target atg mRNA measurements under all of the various conditions tested. However, this would have required a very large number of qPCR tests to be run; we note that in previous assessments of atg RNAi effects on lifespan (listed in Table S1 and Table S6), such tests were rarely performed. Regarding RNAi effects on daf-2 mutants, we note in the text the following: “While mRNA levels after RNAi under the various other conditions tested were not assayed, reduced IIS (including daf2(e1370)) has been shown to intensify the RNAi response (Wang and Ruvkun, 2004), thus lack of effect on lifespan in daf-2(e1370) is unlikely to reflect suppression of mRNA knockdown.” Here we have at least assessed, using N2, the most important issue relating to RNAi efficacy: the differential effects of different RNAi feeding clones on atg mRNA levels.

(5) Incomplete mechanistic exploration. The investigation of why atg-18 knockdown has uniquely strong effects was limited to DAF-16. Given published evidence that atg-18 may regulate HLH-30/TFEB, a master transcriptional regulator of autophagy and lysosomal biogenesis, testing whether atg-18 specifically affects HLH-30 nuclear localisation or activity could have provided valuable mechanistic insight and would distinguish atg-18 from the other genes tested.

We would have readily investigated this. However we learned of the interactions between atg-18 and hlh-30 only in Nov 2025, when one of us (David Gems) bumped into a member of Evandro Fan’s research group at an aging meeting at the Crick Institute in London. Their findings were very interesting for us, as they offered a possible explanation for the seeming idiosyncrasy of atg-18 RNAi effects on lifespan. This subject is currently under investigation by the Fan lab at the University of Oslo, who recently posted a preprint describing the work (Schmauck-Medina et al., 2026).

Reviewer #2 (Public review):

Summary:

This study examines how genes involved in cellular recycling (autophagy) influence lifespan under different experimental conditions. The findings help clarify why previous studies have reported conflicting results about whether blocking autophagy shortens or extends lifespan. The work will be of interest to researchers studying aging and cellular stress responses, particularly those using model organisms.

We thank the reviewer for their helpful remarks. Responding to their comments (particularly those not in the public review) has enabled us to improve it.

Strengths:

The findings are valuable, as they help resolve inconsistencies within a specific subfield of aging research. The evidence presented is solid, as the data broadly support the primary claims of the study. In addition, the discussion is thorough and thoughtfully integrates the findings within the broader context of the field.

Weaknesses:

Additional functional validation would further strengthen the conclusions.

We very much agree. Our original plan for this study was to include autophagic flux assays under different conditions. However, this line of investigation led us to a careful reassessment of reporter-based approaches to measuring autophagic flux in C. elegans, and attempts to improve them. This includes development of an automated, AI-based quantitative image analysis pipeline to improve reproducibility and data interpretation across studies. This investigation is still ongoing, and we are currently preparing a separate manuscript focused specifically on methodological clarity and quantitative assessment of autophagic flux.

Recommendations for the authors:

Reviewing Editor Comments:

To increase the evidence provided by the authors, they should at least address the comments from Reviewer 1 regarding the experimental inconsistencies. We acknowledge that the additional experiments suggested by Reviewer 2 regarding the use of C. elegans mutants and additional methods to assess autophagic flux would likely be a lot of additional work. However, adding results from such experiments would, of course, make the evidence more compelling.

Reviewer #1 (Recommendations for the authors):

Writing and presentation

(1) The abstract discusses results for daf-2 in detail but does not mention the glp-1 findings. Given that a substantial portion of the study addresses glp-1 longevity, including a summary of those results in the abstract would better represent the scope of the work.

Agreed. glp-1 is now referred to in the abstract.

(2) In the abstract, the sentence regarding FUDR effects is placed between statements about daf2, while it is referring to N2 lifespans, which may give the impression that the FUDR results were obtained in a daf-2 background. Consider restructuring this section for clarity.

Agreed. To improve clarity this now reads as follows. “In wild-type C. elegans, FUDR at a high concentration caused knockdown of several atg genes to increase lifespan”

(3) The definition of "robust" used in Figure 4E could be misleading to readers. For instance, bec-1 and atg-4.1 knockdowns in glp-1 at 20°C are classified as robust, yet the actual percent suppression is modest (~8% and ~3%, respectively) and non-consistent in individual replicates. The "robust" designation arises because these knockdowns slightly increased N2 lifespan. Clarifying the definition in the figure legend or text would help readers interpret this correctly.

For glp-1 at 20°C robust suppression is seen with atg-2, atg-18 and bec-1 RNAi, not atg-4.1 RNAi. But regarding bec-1: yes, the suppression is modest. bec-1 RNAi is an unusual case insofar as it meets the <30% definition partly because it caused an increase in N2 lifespan. Under the circumstances, arguably, it makes little sense to view it as an example of robust suppression, and Figure 4E has been altered accordingly, with a note added to the legend as follows. “Note that the bec-1 RNAi effect on glp-1 at 20°C is not classified as robust here even though it reduces lifespan to within <30% of the mean lifespan of N2 under bec-1 RNAi, since the fact that it does so partly reflects an increase in N2 lifespan, rather than a robust life-shortening effect on glp-1.”

To try to improve clarity we have altered the definition of “robust” to read as follows. “R, robust suppression, i.e. knockdown reduces the extended lifespan of daf-2 or glp-1 to within <30% of the mean lifespan of N2 under the same RNAi. This designation (“robust”) indicates a high degree of suppression of the mutant longevity phenotype (see Figure 2, Figure 5 and Fig. S2).”

(4) On page 12, the text should read "9/30 suppresses robustly" (currently appears to contain a numerical error).

Fixed. This now reads “In 8/30 the RNAi effect was robust, i.e. the mutant longevity was largely suppressed.” (Now 8/30 since bec-1/glp-1/20°C is no longer viewed as robust suppression).

(5) In Supplementary Sheet 8, the lifespan data from Hashimoto et al. are presented in a different format than the data from other studies. Standardizing the presentation would improve readability.

This is Supplementary Table 1. The inconsistencies have been ironed out.

Data and calculations:

(6) In Supplementary Table S3, the ΔΔCt values for atg-9 appear to be incorrect. Please verify and correct.

We thank the reviewer for highlighting this error. The ΔΔCt values for atg-9 in Supplementary Table S3 were incorrectly entered; the Fold Change values had been mistakenly placed in that column. This error has now been corrected. Please note that analyses and conclusions reported in the manuscript were based on the correct values.

(7) In Supplementary Table S4, the standard deviation values do not match my independent calculations. Please double-check these values.

This is correct: there was an error in the standard deviation (SD) values. We thank the reviewer for their diligence. We have updated both the SD and SEM (standard error of mean) columns in Table S4. The mean ΔΔCt values remain unchanged, as do the conclusions from analyses and statistical tests.

(8) In Supplementary Table S7 (kanamycin experiment), there appear to be several errors in the percent change calculations. Additionally, the statement "In the absence of Kan, atg-13 RNAi caused a slight reduction in lifespan" is not supported by the combined data, which actually shows a slight increase. Given that the reported changes are subtle but statistically significant, it would be prudent to re-verify the p-value calculations as well before drawing conclusions from this experiment.

The calculation for trial 1 atg-13 (-Kan) as a percentage of control (L4440 Kan) has been corrected so that the mean lifespan of the knockdown is divided by the mean lifespan of the control. In the previous version, the ratio was inadvertently calculated in the opposite orientation. All other values remain unchanged.

Responding further to this point, to strengthen the data here we have also conducted an additional trial, and Figure 3D and Table S7 have been updated accordingly. The more robust data still supports the conclusion that E. coli infection does not mask a life extending effect of atg-13 RNAi. However, in the new, summed data, atg-13 RNAi on no Kan does not shorten lifespan at all, in contrast to our previous trials (conducted several years earlier), but consistent with several other instances of variability in the study. Moreover, the modest life-shortening effect of atg-13 RNAi on Kan (-6.7%) is now statistically significant. The manuscript has been updated accordingly.

Experimental interpretation

(9) On page 9, the authors report testing N2 and daf-2(e1370) lifespan at 15°C and 20°C, but only the daf-2 results are discussed in the text. The N2 results at 15°C appear in Table S5 but are never addressed. Notably, bec-1 knockdown significantly suppressed N2 lifespan at 20°C in Table S2 but appears to significantly extend it at both 15°C and 20°C in Table S5. These discrepancies should be discussed.

The issue of inter-trial variability is discussed in our response to point 1 in the public review. More specifically: here it may be significant that the trials listed in Table S5 were performed several years after those in Table S2. The discrepancy is now noted and discussed as follows. “In these trials bec-1 RNAi also modestly increased N2 lifespan at both temperatures (Table S5), surprisingly given that in previous trials (performed several years earlier) bec-1 RNAi shortened N2 lifespan (Table S2). The reason for this discrepancy is unknown.”

(10) Regarding the CPH analysis of bec-1 in daf-2(e1368) (Figure 1), the authors state that bec-1 knockdown does not have a significantly greater effect in daf-2(e1368) relative to N2, and then note that this is consistent with the earlier observation by Hansen et al. that bec-1 shortens daf-2 lifespan without affecting N2. However, in the cumulative dataset, bec-1 does significantly suppress N2 lifespan. A more precise statement here would prevent readers from drawing an incorrect conclusion.

The point here is that our study and the Hansen et al study both point to atg RNAi suppression of daf-2 Age being limited to class 1 mutants, not that there are no effects on N2. To try to improve clarity we have rephrased as follows. “These findings are broadly consistent with the earlier observation that bec-1 and vps-34 RNAi shortened the lifespan of the daf-2(mu150) class 1 mutant but not of N2 at 20°C (CPH analysis not performed) (Hansen et al., 2008).”

(11) The reference to Hashimoto et al. (2009) on page 10 states that atg-9 and atg-13 RNAi increased lifespan, but that study does not include data for atg-13. The lifespan extension reported by Hashimoto et al. was for atg-7, atg-9, bec-1, and unc-51. Please correct this citation.

Done. It now reads “where atg-9 (and also atg-7, bec-1 and unc-51) RNAi increased lifespan”.

(12) In a previous publication from this group, atg-2 and atg-13 knockdown with 15 µM FUDR led to significant lifespan extension, whereas in the current study, the same treatments significantly suppressed lifespan. Although this discrepancy is briefly mentioned in the Discussion, a more thorough discussion of possible explanations would strengthen the manuscript's value as a reference dataset for future studies.

A more detailed discussion has been added, as follows. “Regarding the causes of variability between results of ostensibly identical tests performed under ostensibly identical conditions: one clue is provided by a study comparing results of lifespan assays performed across three sites under similar conditions. This revealed that inter-trial variation occurred mainly at each site over time, rather than between sites (Lucanic et al., 2017). One possibility is that this reflects batch variation in media components, such as the BactoPeptone constituent of nematode growth medium (Petrascheck, 2014).”

(13) The N2 mean lifespan on GFP RNAi with 0 µM FUDR at 20°C is approximately 15 days, whereas the N2 lifespan at 20°C in Table S2 is approximately 20 days. While inter-experiment variability is expected, a difference of this magnitude warrants acknowledgement, as it could influence the interpretation of subsequent comparisons.

We have now acknowledged this in the legend to Figure 3, as follows. “We note that in (A) the lifespan of the gfp RNAi control is somewhat lower than in other experiments (mean 14.96 days, Table S6); see Discussion for consideration of possible reasons for inter-trial variability.”

(14) Regarding the glp-1 experiments (page 11 and Table S9), the N2 lifespan values in these experiments differ from earlier N2 results at 20°C for several knockdowns (e.g., bec-1 knockdown appears to increase N2 lifespan in Figure 4). Additionally, the glp-1 lifespan results are not consistent between the two replicates for most genes except atg-2 and atg-18. The developmental shift (raised at 25°C, then moved to 20°C to obtain the glp-1 phenotype) could plausibly account for some of this variation compared to animals raised continuously at 20°C. If so, this should be explicitly discussed.

Agreed. The following has been added to the Figure 4 legend. “That bec-1 RNAi increases N2 lifespan in (A) (+16.7%, p < 0.0001) but not (B) could imply an interaction with temperature during development, or merely variability of atg RNAi effects (see Discussion).”

(15) At 25°C, N2 lifespan shows significant suppression upon atg-13 and atg-18 knockdown in Table S9, while these same effects were non-significant in Table S2. These and other interexperiment discrepancies should be noted and, if there are identifiable experimental differences, those should be specified.

This discrepancy has now been noted on page 9, immediately after the description of the data in Table S2, as follows. “(although in later tests at 25°C, life-shortening effects of atg-13 and atg-18 RNAi were seen; Table S9)”

(16) If autophagy is already regulated by heat stress at 25°C, this could explain the diminished effects of autophagy gene knockdown at higher temperatures. Measuring autophagy gene expression by qPCR across different temperatures and genetic backgrounds could provide useful mechanistic insight.

Agreed. However, more informative will be to measure effects of temperature and genotype on autophagy more directly, using fluorescent reporters of autophagic flux. We are addressing this as part of an ongoing study using improved and fully validated autophagic flux measurement methodologies (please see our response to reviewer 2, public review).

(17) The authors report that atg-18 knockdown upregulates other autophagy genes (supplementary data). This is intriguing given that atg-18 shows the strongest phenotype. Whether this reflects a compensatory mechanism and why it does not rescue the lifespan suppression deserves further discussion.

On reflection we decided to remove from the manuscript the data relating to effects of atg-18 RNAi on mRNA levels of other atg genes due to concerns about data quality.

(18) Regarding the FUDR and infection hypothesis: the logic that reduced bacterial infection upon FUDR treatment explains the loss of lifespan suppression is reasonable, but it does not account for why atg-13 knockdown actively extends lifespan in the presence of FUDR. This point could benefit from further discussion. In the kanamycin experiment, they further see that life-shortening effects of atg RNAi are not solely attributable to infection, but the question of lifespan extension remains unanswered.

Good point. We have addressed this as follows. “We also conclude the increase in lifespan upon atg-13 RNAi in the presence of 800 μM FUDR (Figure 3C) is not attributable to suppression of E. coli infection, but rather to some other, unidentified mechanism.”

(19) The timing of when lifespan assays are initiated relative to other experimental treatments is another potential source of variability (as seen from earlier reported data) that could be acknowledged as a consideration for future studies.

Good point. We have added the following to the section of the discussion about tackling condition dependency issues. “Another factor to take into account is the apparent tendency of results of C. elegans lifespan assays to vary over time (Lucanic et al., 2017).”

Reviewer #2 (Recommendations for the authors):

Major Comments:

(1) Assessment of autophagic activity

The authors demonstrate by qPCR that feeding RNAi reduces mRNA levels of autophagy-related genes. However, reduced transcript levels do not necessarily confirm functional inhibition of autophagy. Incorporating established assays of autophagic flux, such as Western blot analysis of lipidated ATG-8 (LGG-1/ATG-8-II) or validated fluorescence-based reporters, would substantially strengthen the mechanistic conclusions.

We agree with the reviewer that direct assessment of autophagic flux would provide additional functional insight. We are currently performing complementary assays to directly assess autophagic flux using reporter-based approaches, with a focus on standardizing reporter-based measurements and developing a quantitative image analysis pipeline to improve reproducibility and interpretation across studies. These analyses will be presented in a separate manuscript focused specifically on methodological clarity and quantitative assessment of autophagic flux.

(2) Use of genetic mutants

Several C. elegans loss-of-function mutants for autophagy genes are available. Validation of key findings using selected genetic mutants, where feasible, would provide complementary evidence and enhance confidence in the RNAi-based results.

In principle is this a good idea. In practice many loss-of-function mutants in core autophagy genes exhibit developmental defects or impaired viability, which can confound interpretation in aging studies. As our aim was to examine the effects of autophagy gene perturbation specifically during adulthood, RNAi provided a practical approach that allowed post-developmental knockdown while minimising disruption of normal development.

Minor Comments:

(1) Context-dependent effects of autophagy

The findings are conceptually consistent with prior work demonstrating dual roles of autophagy in C. elegans survival during starvation, where physiological levels promote survival but insufficient or excessive autophagy contributes to mortality (Kang et al., Genes & Development, 2007). Including a discussion of this study would help frame the present results within a broader biological context.

Good idea to cite this study, and we have now done so in the introduction, as follows. “It is by now clear that autophagy can enhance as well as inhibit the development of pathologies in C. elegans, including senescent ones (Kang et al., 2007).”

(2) Relevance beyond C. elegans

It would be helpful for the authors to clarify whether similar context-dependent effects of autophagy on lifespan have been reported in other organisms. Briefly referencing comparable findings in additional model systems would broaden the relevance and impact of the study.

This is a good idea, but our search for similar cases in other model organisms failed to identify clear examples. Perhaps more to the point here is that context dependent effects and, perhaps, condition selection bias, are a serious issue in scientific research in general. To emphasize this, the following has been added as the last line of the discussion. “More widely, condition dependency and conditional selection bias risk diminishing the reliability of research findings in many scientific disciplines.”

Hashimoto, Y., Ookuma, S. and Nishida, E., 2009. Lifespan extension by suppression of autophagy genes in Caenorhabditis elegans. Genes Cells. 14, 717-726.

Lucanic, M., Plummer, W., Chen, E., Harke, J., Foulger, A., Onken, B., Coleman-Hulbert, A., Dumas, K., Guo, S., Johnson, E., Bhaumik, D., Xue, J., Crist, A., Presley, M., Harinath, G., Sedore, C., Chamoli, M., Kamat, S., Chen, M., Angeli, S., Chang, C., Willis, J., Edgar, D., Royal, M., Chao, E., Patel, S., Garrett, T., Ibanez-Ventoso, C., Hope, J., Kish, J., Guo, M., Lithgow, G., Driscoll, M. and Phillips, P., 2017. Impact of genetic background and experimental reproducibility on identifying chemical compounds with robust longevity effects. Nat Commun. 8, 14256.

Minnerly, J., Zhang, J., Parker, T., Kaul, T. and Jia, K., 2017. The cell non-autonomous function of ATG-18 is essential for neuroendocrine regulation of Caenorhabditis elegans lifespan. PLoS Genet. 13, e1006764.

Schmauck-Medina, T., Anisimov, A., Meyer, D.H., Hu, Y., Huang, Z., Wu, Y., Taylor, S., Takla, M., MacArthur, M.R., Mitchell, S.J., Ai, R., Simonsen, A., Jensen, V., Labbadia, J., Shen, H.-M., Hansen, M., Schumacher, B., Rubinsztein, D., Lautrup, S., Lu, G. and Fang, E.F., 2026. ATG-18/WIPI2 drives longevity in an HLH-30/TFEB-dependent manner. bioRxiv.

Wang, D. and Ruvkun, G., 2004. Regulation of Caenorhabditis elegans RNA interference by the daf-2 insulin stress and longevity signaling pathway. Cold Spring Harb Symp Quant Biol. 69, 429-31.

Wilhelm, T., Byrne, J., Medina, R., Geisinger, J., Hajduskova, M., Tursun, B. and Richly, H., 2017. Neuronal inhibition of the autophagy nucleation complex extends life span in postreproductive C. elegans Genes and Development. 31, 1561–1572.

Yang, Y., Arnold, M.L., Lange, C.M., Sun, L.H., Broussalian, M., Doroodian, S., Ebata, H., Choy, E.H., Poon, K., Moreno, T.M., Singh, A., Driscoll, M., Kumsta, C. and Hansen, M., 2024. Autophagy protein ATG-16.2 and its WD40 domain mediate the beneficial effects of inhibiting early-acting autophagy genes in C. elegans neurons. Nat Aging. 4, 198-212.

Version 2

Abstract

Autophagy is proposed to protect against aging by clearing damaged cellular constituents. In line with this several life-extending interventions in model organisms show some degree of autophagy dependence. In C. elegans, inhibiting autophagy can shorten, lengthen or have no effect on lifespan. Differences between published findings likely reflect variability in experimental conditions. Here we investigate the condition dependence of effects on lifespan of RNA-mediated interference (RNAi) knockdown of autophagy pathway components. Effects on interventions causing a strong Age (increased lifespan) phenotype were examined: mainly mutation of daf-2 (insulin/IGF-1 receptor), but also suppression of germline development by mutation of glp-1. Factors varied included daf-2 mutant allele class, atg gene, temperature and presence of 5-fluoro-2’-deoxyuridine (FUDR). Effects on lifespan of atg RNAi proved to be highly condition dependent. Notably, for most atg genes tested lifespan was not usually reduced more in the long-lived mutant than in the wild-type control. Greater suppression was seen at 20°C for certain atg genes with daf-2(e1368) but not daf-2(e1370). At 25°C, little reduction in lifespan was seen. However, atg-18 knockdown behaved differently, suppressing daf-2 Age under all conditions, suggesting possible pleiotropic action. In wild-type C. elegans, FUDR at a high concentration caused knockdown of several atg genes to increase lifespan. Thus, depending on experimental conditions, atg knockdown can increase, decrease or have no effect on daf-2 Age. Condition dependent effects were also seen with respect to glp-1 Age. The lack of suppression of daf-2 Age by atg RNAi under most conditions questions the importance of autophagy for this phenotype. Moreover, condition dependence of effects creates a risk of possible condition selection bias.

Introduction

A long-standing theory is that senescence (aging) is largely a consequence of the accumulation of random molecular damage caused by, among other things, reactive oxygen species (Beckman and Ames, 1998; Harman, 1956; Murphy, 2023; Shore and Ruvkun, 2013; Szilard, 1959). This view predicts that mechanisms of somatic maintenance, particularly those that prevent accumulation of damaged cellular constituents, counteract the aging process. One somatic maintenance function viewed as a potential longevity-assurance process is autophagy (specifically macroautophagy), which effects lysosome-dependent degradation of cellular constituents, including damaged matter (Aman et al., 2021).

The possible anti-aging role of autophagy has been extensively tested in the short-lived nematode Caenorhabditis elegans, and supporting evidence found with respect to several life-extending interventions, including reduced insulin/IGF-1 signaling (IIS), reduced germline signaling, and dietary restriction (Hansen et al., 2018). These studies were principally performed in the 2000s; however, during the same period, falsification tests of the molecular damage theory, particularly that relating to oxidative damage, led to growing uncertainty about its validity (Gems and Doonan, 2009; Perez et al., 2009; Shields et al., 2021). Meanwhile, an alternative theoretical framework emerged, based on the evolutionary theory of aging (Arnold and Rose, 2023; Williams, 1957), arguing that senescence is largely the consequence of genetically-determined, programmatic mechanisms (Blagosklonny, 2006; de Magalhães and Church, 2005; Gems, 2022; Maklakov and Chapman, 2019), and very much so in C. elegans (Gems and de la Guardia, 2013; Gems et al., 2021; Pires da Silva et al., 2024).

One form of programmatic aging involves costly programs: genetically-determined processes that consume or degrade somatic tissues as a by-product of wider, fitness-promoting processes (Gems and Kern, 2024; Gems et al., 2021). In some cases this involves biomass repurposing, where biomass of one tissue is broken down by autophagic processes to release molecular constituents to support functions in another. This occurs to a high degree in semelparous organisms in the process of reproductive death (suicidal reproductive effort), as in many monocarpic plants, and semelparous fish such as Pacific salmon (Gems et al., 2021).

Several lines of evidence support the hypothesis that reproductive death occurs in C. elegans hermaphrodites (Gems et al., 2021; Kern et al., 2023). This includes a putative costly program in which intestinal biomass is broken down and repurposed to support production of yolk, that is then vented to support larval growth, leading to intestinal atrophy (a senescent pathology) (Ezcurra et al., 2018; Kern et al., 2021; Sornda et al., 2019). Notably, such biomass repurposing is supported by autophagy, as evidenced by deceleration of intestinal atrophy and yolk pool formation when autophagy is inhibited (Benedetto and Gems, 2019; Ezcurra et al., 2018). Thus, in this particular context autophagy appears to be promoting rather than preventing senescence. It is by now clear that autophagy can enhance as well as inhibit the development of pathologies in C. elegans, including senescent ones (Kang et al., 2007).

These developments warrant a careful reassessment of the evidence that autophagy is protective against aging in C. elegans. In this study we reexamine the question of whether the longevity of daf-2 insulin/IGF-1 receptor mutants is autophagy dependent. Here the principal form of past evidence involves epistasis tests, where effects on lifespan of reduction of function of genes encoding proteins effecting autophagy is compared in the wild type (N2) and daf-2 mutants. Findings from 9 previous studies involving 46 epistasis experiments are summarised in Table S1. Although it is widely believed that autophagy is essential for daf-2 mutant longevity (Meléndez et al., 2003), scrutiny of the results of past tests raises doubts about the strength of this claim.

Careful consideration of these prior studies identifies six distinct limitations, as follows.

(1) A life-shortening effect of inhibition of autophagy does not necessarily indicate its role in the normal aging process, or in daf-2 mutant longevity. (2) If autophagy is inhibited during development as well as adulthood, a life-shortening effect could result from disruption of normal development. (3) The claim that daf-2 longevity is autophagy dependent requires evidence that the life-shortening effect in daf-2 is significantly greater than in wild type, e.g. by using Cox proportional hazard analysis, and this is rarely performed. (4) If effects of reducing autophagy are condition dependent, this introduces a potential bias: a risk that investigators might unwittingly select conditions where the data generated supports a role of autophagy in longevity - what may be referred to as condition selection bias. Potential determinative conditions that have varied across studies include choice of autophagy-determining gene to inhibit, of daf-2 mutant allele, and ambient temperature. (5) 5-fluoro-2’-deoxyuridine (FUDR) is sometimes used to facilitate C. elegans lifespan assays by preventing progeny hatching; this could potentially affect test outcomes, as shown in other contexts (Aitlhadj and Sturzenbaum, 2010; Anderson et al., 2016; Burnaevskiy et al., 2018; Van Raamsdonk and Hekimi, 2011; Zhao et al., 2019). (6) A final, straightforward issue is whether a given finding proves to be reproducible in subsequent reports under, seemingly, the same conditions.

Of 46 prior tests (Table S1) only 4 present clear evidence that reducing autophagy shortens lifespan more in daf-2 than in the wild type. Regarding one of these four instances, the effect of bec-1 RNAi on daf-2(e1370) at 15°C (Meléndez et al., 2003), a subsequent study did not replicate this finding (Hars et al., 2007). In two other cases the weaker daf-2(mu150) allele was used (Hansen et al., 2008; Patel et al., 2008). In two further studies where large reductions in daf-2 lifespan were observed (Chang et al., 2017; Minnerly et al., 2017) knockdown of atg-18 was used. In a 2009 study where effects of knockdown of diverse autophagy genes was tested, atg-18 was one of only 2/14 genes where adult-limited RNAi significantly reduced lifespan in the wild type (Hashimoto et al., 2009), suggesting possible atg-18 idiosyncrasy. The 2009 study includes more than half of all of the prior tests (28/46); strikingly, in only 1/14 genes (atg-4.1) did adult-limited RNAi significantly reduce daf-2(e1370) lifespan, while for 3/14 genes (atg-9, bec-1 and unc-51) it increased daf-2 lifespan (Hashimoto et al., 2009). One possible reason for the lack of observed life-shortening effects given autophagy knockdown is its use of FUDR (high concentration), unlike other studies.

Here we assess the condition-dependence of the effects of RNAi knockdown of autophagy genes on C. elegans longevity. To this end we have tested the effects of RNAi of six genes in the autophagy pathway on longevity in two different daf-2 mutants, and a glp-1 germlineless mutant, at two temperatures. We have also assessed effects of FUDR and prevention of bacterial infection. Our results provide a robust foundation of evidence relating to possible autophagy dependence of daf-2 and glp-1 longevity. They suggest a weak and highly-condition dependent contribution of autophagy to daf-2 and glp-1 Age. This suggests that prevention of damage accumulation by autophagy is, at most, a minor determinant of daf-2 longevity.

Results

Effects of atg RNAi on daf-2 Age vary with daf-2 allele

We first tested whether effects on lifespan of inhibiting autophagy depend upon daf-2 allele severity. Two daf-2 mutants were examined. daf-2(e1368) is a class 1 (less pleiotropic) mutant where adults show normal behavior at both 20°C and 25°C, while daf-2(e1370) is a class 2 (more pleiotropic) mutant where at 25°C adults become paralyzed and cease feeding (Gems et al., 1998). Knockdown of autophagy was performed by RNAi from the L4 (late larval) stage, to exclude possible confounding, life-shortening disruption of normal development. RNAi was performed for six genes at several stages of the autophagy process: initiation (atg-13), membrane nucleation (atg-9, bec-1), phagophore formation (atg-2, atg-18) and elongation (atg-4.1) (Figure 1A). 5/6 genes were those examined in our previous study (Ezcurra et al., 2018), with bec-1 added because it was the subject of several earlier studies (Table S1).

Effects of atg RNAi on longevity are highly variable and condition dependent.

(A) The autophagy pathway, and genes tested in this study. (B-D) Effects at 20°C. (B) Effects on N2 (wild type). (C) Effects on daf-2(e1368). (D) Effects on daf-2(e1370). (E, F) Effects at 25°C. (E) Effects on N2. (F) Effects on daf-2(e1368). (G) Effects of bec-1 RNAi on daf-2(e1370) at 15°C. (B-G summed data, N = 2; for individual trial data and statistical comparisons, see Table S2, S5.

A methodological note: for tests of effects of a given intervention on C. elegans lifespan an often-applied standard is to include 3 biological replicates. This is true of several recent studies where the effect of knockdown of a single atg gene on daf-2 longevity was studied (Minnerly et al., 2017; Wilhelm et al., 2017; Yang et al., 2024). However, given that the present condition dependence study effectively performs this test in 18 different ways, involving RNAi of 6 atg genes, 2 daf-2 mutants and 2 temperatures, N = 2 biological replicates were judged to be sufficient to draw robust conclusions; similarly, an earlier study of RNAi 14 atg genes under two conditions used 2-3 biological replicates (Hashimoto et al., 2009); for an overview of the number of replicate trials performed in previous studies, see Table S1.

Trials were performed at 20°C, and used gfp RNAi as a negative control (with induction of an RNAi response). In wild-type C. elegans (N2), statistically significant reductions were sometimes seen: atg-9, atg-13 and, particularly, atg-18 RNAi shortened lifespan in both trials (summed data, N = 2, means: -13.7%, p < 0.0001, -12.4%, p = 0.0002, -31%, p < 0.0001, respectively; log rank test) (Figure 1B, Table S2; for all raw mortality data, see Supplementary Dataset 1). For survival curves comparing RNAi effects of individual atg genes on the three genotypes, see Figure 2.

Degree of suppression of daf-2 longevity by atg gene RNAi differs greatly between daf-2 alleles and atg RNAi treatments (20°C), cf. Figure 1B-D (same data).

Summed data, N = 2; for individual trials and statistical comparisons, see Table S2. All trials were performed in parallel, i.e. all lifespans (summed or in each trial) are directly comparable.

RNAi of all six atg genes consistently reduced lifespan in the daf-2(e1368) mutant, with effects ranging from -10.5% (p = 0.0009) for atg-4.1 to -50.4% (p < 0.0001) for atg-18 (Figure 1C, Table S2). To assess whether atg RNAi reduced lifespan more in daf-2 mutants than in the wild type, Cox proportional hazard (CPH) analysis was used. This showed a significantly greater effect in daf-2(e1368) for 4/6 genes (exceptions: atg-4.1, bec-1) (Table S2). These findings are broadly consistent with the earlier observation that bec-1 and vps-34 RNAi shortened the lifespan of the daf-2(mu150) class 1 mutant but not of N2 at 20°C (CPH analysis not performed) (Hansen et al., 2008).

To test whether atg RNAi is able to fully suppress daf-2(e1368) Age, the lifespans of N2 and daf-2 subjected to atg RNAi were compared. Under all six atg RNAi conditions, daf-2(e1368) still significantly increased mean lifespan, from +15.7% (p = 0.0031) for atg-13 to +60.9% (p < 0.0001) for atg-4.1 (summed data, 20°C, Table S2). This could imply either that daf-2(e1368) Age is not fully autophagy dependent, or that it is but autophagy is not fully suppressed by the RNAi treatments used.

In daf-2(e1370), atg RNAi had far more modest effects on lifespan than in daf-2(e1368). Summed data showed significant reductions in lifespan resulting from RNAi of atg-9, atg-13 and atg-18 only, with the latter again causing the largest reduction: -8.3% (p = 0.012), -17.8% (p = 0.0034), and -29.2% (p < 0.0001), respectively (Figure 1D, Table S2). However, effects were in no instance significantly greater than in N2 (CPH analysis, Table S2), thus failing to provide evidence that e1370 longevity is mediated by autophagy. Taken together, these results suggest that at 20°C autophagy contributes to longevity in weaker but not stronger daf-2 mutants, possibly in class 1 but not class 2 mutants.

RNAi effects varied between atg genes, with atg-2 and atg-4.1 RNAi effects weaker, and atg-18 RNAi effects generally stronger than the rest. One possible cause of this variability is differences between the plasmid vectors used for RNAi in terms of how efficiently they destroy their target mRNA. To investigate this possibility, RNAi of each atg gene was performed in N2 animals, and transcript levels measured by RT–qPCR and analyzed using the ΔΔCt method, normalized to a gfp RNAi control, with fold changes calculated as 2^−ΔΔCt (N = 4 independent trials). Effects of RNAi on mRNA level varied greatly, from no detected reduction in atg-2 to an 86% reduction in atg-18 (Fig. S1A; Table S3; Supplementary dataset 2). Pairwise comparisons of ΔΔCt values showed that, atg-2 aside, there were no significant differences between RNAi treatment effects on mRNA levels, apart from a greater effect of atg-18 RNAi when compared to either atg-4.1 and atg-9 RNAi (Fig. S1B).

Comparing effects of RNAi on mRNA levels and lifespan, the only clear correspondence between the former and the latter involved atg-2 and atg-4.1 RNAi, which in N2 had no significant effect on either (Figure 1B, Fig. S1A; Table S2). However for the remaining 4 genes, there was no clear correspondence, though atg-18 mRNA levels appear to be the lowest, in line with its greater effect on lifespan (Figure 1B-D, Fig. S1A; Table S2, Table S3). These results imply that RNAi efficacy and perhaps also gene-specific issues contribute to variation in atg RNAi effects on lifespan (discussed further below). While mRNA levels after RNAi under the various other conditions tested were not assayed, reduced IIS (including daf-2(e1370)) has been shown to intensify the RNAi response (Wang and Ruvkun, 2004), thus lack of effect on lifespan in daf-2(e1370) is unlikely to reflect suppression of mRNA knockdown.

Effect of atg RNAi on daf-2 Age is temperature dependent

Results of a previous study performed at 25°C appear to show no greater reduction in lifespan in daf-2(e1370) compared to daf-2(+) after autophagy knockdown, even when using the atg-18(gk378) deletion mutation (Toth et al., 2008) (Table S1), suggesting possible temperature dependence of atg RNAi effects. To explore this further, in parallel to tests at 20°C, we also compared effects of atg RNAi on lifespan in N2 and daf-2(e1368) at 25°C. Effects on daf-2(e1370) were not tested, partly because this mutant ceases feeding at 25°C (Gems et al., 1998) which would be expected to interfere with RNAi by feeding.

At 25°C, atg RNAi did not shorten N2 lifespan for any of the six genes tested, not even atg-18 (summed data; Figure 1E, Table S2), i.e. culture at 25°C suppressed the life-shortening effect of atg RNAi in N2 (although in later tests at 25°C, life-shortening effects of atg-13 and atg-18 RNAi were seen; Table S9). Also notable is that the increases in lifespan with atg-2 and atg-13 RNAi at 25°C, described in our earlier study (Ezcurra et al., 2018), were not reproduced (discussed below).

At 25°C only atg-18 RNAi significantly reduced daf-2(e1368) mean lifespan, by 15.5% (p < 0.0001) (summed data, Figure 1F, Table S2), a reduction that was significantly greater than in N2 (p = 0.0012, CPH, summed data only; Table S2). For survival curves comparing RNAi effects of individual atg genes on the two genotypes, see Fig. S2. In one case, atg-9, RNAi slightly increased daf-2 lifespan (+9.8%, p = 0.011, summed data).

The initial tests showing that bec-1 RNAi reduces daf-2(e1370) lifespan were performed at 15°C (Meléndez et al., 2003); moreover, a greater N2 life-shortening effect of bec-1 RNAi at 15°C than 20°C has been reported (Chen et al., 2019). Taken together with the weaker RNAi effects at 25°C observed here, this suggested that stronger effects might be seen at 15°C. To explore this we compared effects of bec-1 RNAi from L4 on N2 and daf-2(e1370) at 15°C and 20°C (N = 2). However, no suppression of daf-2(e1370) Age by bec-1 RNAi was seen at either temperature (Figure 1G; Table S5), consistent with findings of an earlier study performed at 15°C (Hars et al., 2007) (Table S1). In these trials bec-1 RNAi also modestly increased N2 lifespan at both temperatures (Table S5), surprisingly given that in previous trials (performed several years earlier) bec-1 RNAi shortened N2 lifespan (Table S2). The reason for this discrepancy is unknown.

FUDR can alter the effect of atg RNAi on lifespan

Since the 1980s FUDR, a thymidylate synthase inhibitor and anti-cancer drug, has sometimes been added to C. elegans survival trials to prevent progeny production (Gandhi et al., 1980; Mitchell et al., 1979). Notably, two previous reports that observed increases in C. elegans lifespan given atg RNAi employed FUDR. Our own study saw increases in N2 lifespan after atg-2 and atg-13 RNAi from L4 with 15 μM FUDR (Ezcurra et al., 2018). Another study that saw increases in N2 lifespan given adult-limited atg-7, atg-9, bec-1 and unc-51 RNAi used FUDR at a higher concentration, 800 μM (Hashimoto et al., 2009) (E. Nishida, personal communication).

To test for FUDR-dependent effects, we compared the impact of atg gene RNAi (atg-2, atg-4.1, atg-9, atg-13, atg-18 and bec-1) on N2 lifespan at 20°C with 0, 15 or 800 μM FUDR (N = 2). With 0 or 15 μM FUDR, only shortening of lifespan was seen, and FUDR did not significantly alter the effects of RNAi (CPH, Figure 3A,B, Table S6).

FUDR but not infection alters outcome on atg gene RNAi (20°C).

(A-C) Effects of 0 μM, 15 μM and 800 μM FUDR. We note that in (A) the lifespan of the gfp RNAi control is somewhat lower than in other experiments (mean 14.96 days, Table S6); see Discussion for consideration of possible reasons for inter-trial variability. (D) No alteration by kanamycin of atg-13 RNAi effect on N2 lifespan. Summed data, N ≥ 2; for individual trials and statistical comparisons, see Table S6, S7.

Addition of 800 μM FUDR increased the lifespan of the gfp RNAi negative control by 61.7% (p < 0.0001, Table S6), perhaps due to prevention of bacterial proliferation, which can otherwise shorten C. elegans lifespan (Garigan et al., 2002; Gems and Riddle, 2000). In the presence of 800 μM FUDR, atg-9 and atg-13 RNAi increased lifespan, by +12.5% (p = 0.014) and +8.7% (p = 0.031), respectively (summed data; Figure 3C, Table S6). Moreover, the life-shortening effect of bec-1 RNAi, seen with 0 or 15 μM FUDR, was absent. Again, atg-18 RNAi robustly reduced lifespan under all three conditions (Figure 3C, Table S6). The results using 800 μM FUDR are broadly in line with those of Hashimoto et al. (2009), where atg-9 (and also atg-7, bec-1 and unc-51) RNAi increased lifespan, and atg-18 was one of only 2/14 genes tested where adult-limited RNAi decreased N2 lifespan. This suggests that the increases in lifespan after atg RNAi reported in that study could have reflected its use of 800 μM FUDR (Hashimoto et al., 2009).

atg RNAi shortens lifespan in the absence of bacterial infection

Under standard laboratory culture conditions, C. elegans lifespan is limited by infection by the E. coli food source, such that prevention of bacterial proliferation substantially increases lifespan (Garigan et al., 2002; Gems and Riddle, 2000). The preceding results could imply that 800 μM FUDR suppresses life-shortening effects of atg RNAi by preventing bacterial infection. Xenophagy is generally protective against infection in C. elegans (Balla et al., 2019; Jia et al., 2009). Thus, reduction in lifespan given atg gene knockdown could reflect increased susceptibility to infection.

To probe this hypothesis, we compared effects on N2 lifespan of atg-13 RNAi at 20°C in the absence or presence of the antibiotic kanamycin (Kan), to suppress bacterial infection. In the absence of Kan, atg-13 RNAi did not alter lifespan in these trials (Figure 3D, Table S7), in contrast to the modest reductions in lifespan seen in other trials in this study (Table S2, S6) (for possible reasons for this discrepancy, see Discussion). Application of Kan extended C. elegans lifespan (+34.2%, p < 0.0001, summed data, Table S7), as previously observed (Garigan et al., 2002), and in its presence atg-13 RNAi resulted in a modest decrease in lifespan (-6.7%, p = 0.0092, Figure 3D, Table S7). These results suggest that life-shortening effects of atg RNAi are not solely attributable to increased susceptibility to E. coli infection. Moreover, they do not indicate marked condition dependency in atg RNAi effects on lifespan with respect to E. coli proliferative status. We also conclude the increase in lifespan upon atg-13 RNAi in the presence of 800 μM FUDR (Figure 3C) is not attributable to suppression of E. coli infection, but rather to some other, unidentified mechanism.

atg-18 RNAi robustly suppresses glp-1(e2141) Age

We next explored more widely the reproducibility and condition dependence of the requirement for autophagy of C. elegans longevity. Prevention of germline development by laser surgery or mutation increases lifespan in C. elegans hermaphrodites (Arantes-Oliveira et al., 2002; Hsin and Kenyon, 1999; Pires da Silva et al., 2024). Prior tests for possible autophagy dependence of such longevity have largely used the temperature-sensitive glp-1(e2141) germline proliferation mutant, which is fertile at 15°C but sterile and with greatly reduced germline development at 25°C. A key study reported strong and reproducible suppression of glp-1 Age by RNAi of five autophagy-related genes, including atg-18 and bec-1 (Lapierre et al., 2011) (previous findings summarized in Table S8).

We first tested the effect of atg RNAi on glp-1 longevity with animals raised from L1 until L4 stage at 25°C, and maintained at 20°C thereafter, similar to previous studies (Table S8). Again, effects of atg-2, atg-4.1, atg-9, atg-13, atg-18 and bec-1 RNA were tested. In the RNAi control glp-1 increased mean lifespan by +50.9% (p < 0.0001, summed data, Table S9). glp-1 lifespan was significantly reduced by atg-2, atg-4.1, atg-9, atg-18 and bec-1 RNAi (but not atg-13 RNAi), and effects were greater in glp-1 than N2 in all cases except atg-4.1 (Figure 4A, B, Figure 5, Table S9). However, suppression was only robust (of a large magnitude) for atg-2 and atg-18 RNAi (Figure 4B, Figure 5). Under all six atg RNAi conditions, glp-1 still significantly increased mean lifespan, from +4.12% (p = 0.0003) for atg-18 to +43.8% (p < 0.0001) for atg-4.1 (summed data, 20°C, Table S9). This could imply either that glp-1 Age is not fully autophagy dependent, or that it is but autophagy is not fully suppressed by the RNAi treatments used.

Degree of suppression of glp-1(e2141) longevity by atg gene RNAi differs greatly between genes.

(A, B) 20°C. (A) Effects on N2. (B) Effects on glp-1. (C, D) 25°C. (C) Effects on N2. (D) Effects on glp-1. Summed data, N = 2-5; for individual trials and statistical comparisons, see Table S9. That bec-1 RNAi increases N2 lifespan in (A) (+16.7%, p < 0.0001) but not (B) could imply an interaction with temperature during development, or merely variability of atg RNAi effects (see Discussion). (E) Overview of effects of suppression of Age by RNAi of genes specifying autophagy. Dark blue, reduction of daf-2 or glp-1 Age that is significantly greater than in N2. Light blue, reduction of daf-2 or glp-1 Age that is not significantly greater than in N2. R, robust suppression, i.e. knockdown reduces the extended lifespan of daf-2 or glp-1 to within <30% of the mean lifespan of N2 under the same RNAi. This designation (“robust”) indicates a high degree of suppression of the mutant longevity phenotype (see Figure 2, Figure 5 and Fig. S2). Note that the bec-1 RNAi effect on glp-1 at 20°C is not classified as robust here even though it reduces lifespan to within <30% of the mean lifespan of N2 under bec-1 RNAi, since the fact that it does so partly reflects an increase in N2 lifespan, rather than a robust life-shortening effect on glp-1.

Degree of suppression of glp-1(e2141) longevity by atg gene RNAi differs greatly between genes (20°C), cf. Figure 4A,B (same data).

Summed data, N = 2-5; for individual trials and statistical comparisons, see Table S9.

In tests with daf-2(e1368) life-shortening effects of atg RNAi were largely absent at 25°C (apart from atg-18) (Figure 1F, Table S2). We therefore wondered whether temperature might also influence the outcome of atg RNAi treatment in glp-1 mutants. To test this we examined RNAi effects on lifespan at 25°C. Under these conditions, glp-1 lifespan was significantly reduced by only atg-2, atg-4.1, and atg-18 RNAi (Figure 4C, D, Fig. S3, Table S9), and effects were significantly greater in glp-1 than N2 only with atg-2 and atg-18 RNAi (CPH analysis, Table S9). In summary, of the six genes tested only atg-2 and atg-18 RNAi robustly suppressed glp-1 Age, and RNAi effects were weaker at 25°C.

For an overview of the effects of RNAi on daf-2 and glp-1, see Figure 4E. In 11 of the 30 conditions tested RNAi knockdown of autophagy caused a greater reduction in lifespan in the long-lived mutant. In 8/30 the RNAi effect was robust, i.e. the mutant longevity was largely suppressed. This suppression was highly condition dependent, differing according to gene knocked down, temperature, daf-2 allele used, and between daf-2 and glp-1. Notably, only atg-18 RNAi robustly suppressed longevity in both daf-2 and glp-1 mutants.

Why are effects of atg-18 RNAi stronger than those of other atg genes?

The particularly strong life-shortening effects of atg-18 RNAi could reflect either a greater reduction in autophagy, or the presence in addition to its effects on autophagy of pleiotropic effects not directly related to autophagy. Given that longevity due to either daf-2 mutation or germline loss are wholly dependent on the FOXO transcription factor DAF-16 (Hsin and Kenyon, 1999; Kenyon et al., 1993), we wondered whether atg-18 RNAi might inhibit DAF-16. To probe this two approaches were taken. First, we used a constitutive dauer formation assay. High population density and food depletion causes C. elegans larvae to form developmentally arrested dauer larvae (Cassada and Russell, 1975). daf-2 mutants undergo constitutive dauer arrest in a temperature-sensitive manner, and this is fully suppressed by daf-16(-) (Riddle et al., 1981). However, using a sensitive assay (daf-2(m41), 22°C, giving a mix of dauer and non dauers), we detected no reduction in the number of dauers formed given atg gene RNAi (atg-2, atg-13 and atg-18) (Fig. S4A).

Second, we tested a GFP reporter for a gene whose expression is elevated in daf-2 mutants in a daf-16-dependent manner (ftn-1), using strains previously constructed (Ackerman and Gems, 2012). GFP levels were compared in daf-2(m577) and daf-16(mgDf50); daf-2 backgrounds. As expected, GFP levels were higher in daf-2 than in daf-16; daf-2 (Fig. S4B). RNAi did not suppress the daf-2-induced increase of ftn-1::gfp expression (Fig. S4B). These findings argue against a pleiotropic effect of atg-18 on DAF-16 function.

Inhibiting autophagy does not reduce vitellogenin accumulation

Finally, we further investigated the hypothesis that autophagy promotes biomass repurposing in C. elegans. Inhibition of yolk synthesis or of autophagy delays intestinal atrophy and yolk pool accumulation, suggesting that intestinal biomass is repurposed for yolk synthesis (Benedetto and Gems, 2019; Ezcurra et al., 2018; Sornda et al., 2019). In principle, this could involve repurposing into yolk protein (vitellogenin) or yolk lipid. To test the former possibility, wild-type hermaphrodites or fog-2(q71) (feminization of germline) mutant females were subjected to atg RNAi. The fog-2 mutant, which lacks self-sperm and so lays no eggs, was used to avoid possible effects of atg RNAi on fertility, reduction of which can increase vitellogenin levels within nematodes (Sornda et al., 2019).

For none of the six atg genes tested did RNAi detectably reduce yolk protein levels, either in N2 hermaphrodites or fog-2 females (Figure 6). This implies that autophagic machinery, including that in the intestine, does not enhance yolk protein production. Thus, if intestinal biomass repurposing occurs, then it likely supports yolk secretion or yolk lipid production.

Absence of effect of atg RNAi on vitellogenin accumulation.

Fold change of yolk proteins YP170, YP115 and YP88 in N2 and fog-2 normalized to gfp day 1 (N = 3). In no case is vitellogenin level significantly different to that in the gfp RNAi control at any time point (two-way ANOVA, Table S10). For raw data for vitellogenin levels see Supplementary Dataset 3.

Discussion

Overall, the effects of reducing atg gene function described here are ambiguous with respect to the role of autophagy in daf-2 or glp-1 longevity, neither strongly supporting or excluding it. However, they are consistent with a role of autophagy in the longevity of weaker, class 1 daf-2 alleles at lower temperatures (Figure 1C,D,F,G, Table S2). Class 1 allele-limited suppression of daf-2 Age has been seen previously, for example in epistasis tests with daf-12 (encoding a dafachronic acid receptor) (Gems et al., 1998; Larsen et al., 1995) and skn-1 (Nrf2-like transcription factor) (Tullet et al., 2008). This could reflect a role of autophagy in longevity assurance limited to conditions of mild IIS reduction, or sensitivity to differential effects of daf-2 on distinct downstream signaling outputs, such as phosphatidylinositol 3-kinase and Ras signaling (Patel et al., 2008).

atg RNAi suppresses daf-2(e1368) Age at 20°C but not 25°C (Figure 1C,F, Table S2). Given that for hypomorphic daf-2 alleles (such as e1368 and e1370) many mutant traits, including Age, show some degree of temperature sensitivity (Gems et al., 1998; Riddle and Albert, 1997), the absence of suppression at 25°C may reflect the increased severity of the daf-2(e1368) mutant phenotype at this higher temperature. However, life-shortening effects of atg RNAi on N2 are also largely absent at 25°C, suggesting that additional mechanisms may also mediate such temperature sensitivity.

Resolving discrepancies between past studies of autophagy dependence of daf-2 Age

The increasing non-reproducibility of many experimental findings, that has been referred to as the reproducibility crisis, is a problem that particularly afflicts biological and biomedical research (Baker, 2016; Lithgow et al., 2017; Prinz et al., 2011; Ritchie, 2020; Voelkl et al., 2020). A strength of C. elegans as an experimental model is the relative ease with which such discrepancies can be resolved, at least in principle. This is thanks to the use of standardized culture conditions across the C. elegans research community, and of nematode strains based on the same isogenic wild-type strain (N2), plus the relatively low cost and short duration of experiments.

Regarding lifespan assays in particular, possible reasons for discrepant findings include clearly identifiable differences in experimental design, such as use or not of FUDR. Less obvious causes include cryptic variation in wild-type genetic background (Zhao et al., 2019), or subtle differences in culture conditions (e.g. due to batch variation in Bacto Peptone, a constituent of nematode growth medium) (Petrascheck, 2014).

Our findings potentially resolve several discrepancies between earlier studies relating to the possible role of autophagy in daf-2 mutant longevity. Previous studies found that inhibiting autophagy either suppressed or enhanced daf-2 Age, or had no effect (Table S1). One study showing suppression used the weak class 1 allele daf-2(mu150) (Hansen et al., 2008). This is consistent with our observation that, atg-18 aside, atg RNAi can suppress Age in a class 1 but not a class 2 allele (Figure 1C,D Table S2).

Does bec-1 RNAi suppress daf-2(e1370) Age? The initial test suggesting this was performed at 15°C, and employed RNAi by injection of the mothers of assayed individuals (Meléndez et al., 2003). In a subsequent study at 15°C, in which both mothers and adult progeny were exposed to bec-1 RNAi by feeding, daf-2(e1370) lifespan appeared not to be shortened (Hars et al., 2007). Several further trials under different conditions did not observe a life-shortening effect either (Hashimoto et al., 2009; Toth et al., 2008). In the present study, we saw no shortening of daf-2(e1370) lifespan in summed data with adult-limited bec-1 RNAi by feeding, at either 15°C or 20°C (Figure 1D,G, Fig. S3, Table S2, S5). Taken together with earlier evidence, our observations suggest that suppression of daf-2(e1370) Age by bec-1 RNAi is not a readily reproducible finding.

Several studies reported lifespan extension following atg RNAi (Ezcurra et al., 2018; Hashimoto et al., 2009). Here we were able to reproduce this effect for several atg genes by applying high dose FUDR (800 μM), thus recapitulating findings by Hashimoto et al (2009), and potentially accounting for the life span increases seen in that study; it was only subsequent to that study that evidence emerged of the capacity of FUDR to alter effects of interventions that impact lifespan (Aitlhadj and Sturzenbaum, 2010; Anderson et al., 2016; Burnaevskiy et al., 2018; Van Raamsdonk and Hekimi, 2011; Zhao et al., 2019).

Returning to issues with replicating published findings: in an earlier study we observed increases in N2 lifespan given atg-2 and atg-13 RNAi (15 μM FUDR present) (Ezcurra et al., 2018). However, this finding proved not to be robust to replication (Figure 3B, Table S6), perhaps reflecting inherent effect variability, condition dependence of an unidentified kind, or a combination of the two. This again illustrates the value, as with the effects of bec-1 RNAi on daf-2(e1370) Age, of repeated verification of effects of interventions on C. elegans lifespan.

Regarding the causes of variability between results of ostensibly identical tests performed under ostensibly identical conditions: one clue is provided by a study comparing results of lifespan assays performed across three sites under similar conditions. This revealed that inter-trial variation occurred mainly at each site over time, rather than between sites (Lucanic et al., 2017). One possibility is that this reflects batch variation in media components, such as the BactoPeptone constituent of nematode growth medium (Petrascheck, 2014).

As with daf-2, our tests with the glp-1 germline proliferation mutant did not clearly support the view that mutant longevity is autophagy dependent. However, consistent with an earlier study (Lapierre et al., 2011), we observed that atg-18 and bec-1 RNAi reduce glp-1 lifespan more than in wild type at 20°C.

Why does atg-18 RNAi more strongly suppress daf-2 Age?

Of the six autophagy-determining genes tested here, RNAi of atg-18 showed a greater capacity to suppress both daf-2 and glp-1 Age than the other five. atg-18 appears to have replaced bec-1 as the gene of choice for autophagy-related epistasis studies in C. elegans (Chang et al., 2017; Minnerly et al., 2017). Notably, atg-18 was one of only 2/14 autophagy genes tested where RNAi during development caused a high level (>50%) of larval growth arrest or lethality (Hashimoto et al., 2009). The more marked effects of atg-18 RNAi could reflect either greater inhibition of autophagy, or pleiotropy in which processes other than autophagy are altered.

Regarding pleiotropy, several proteins in the canonical autophagy pathway have recently been found to participate in other processes. For example, ATG8 (LGG-1 in C. elegans) functions in trafficking of single-membrane organelles (Nieto-Torres et al., 2021), and C. elegans ATG-16.2 contributes to neuronal exopher formation via its WD40 domain (Yang et al., 2024). In an as yet unexplained instance of atg gene idiosyncrasy, mutation of atg-16.2, atg-18 or bec-1 retards the cell cycle in C. elegans germline cells, while that of atg-7 does not (Ames et al., 2017).

Concerning atg-18, several previous studies suggest possible pleiotropy. ATG-18 is a predicted WIPI (WD repeat protein Interacting with PhosphoInositides) family member. In humans there are four WIPI proteins, WIPI1 - WIPI4. C. elegans atg-18 is more closely related to WIPI1/WIPI2, while epg-6 (Ectopic P Granules 6) more closely resembles WIPI3/WIPI4 (Lu et al., 2011). Notably, atg-18 and epg-6 deletion mutations appear to cause similar reductions in levels of autophagy, yet only atg-18 shortens lifespan (Takacs et al., 2019).

Next, C. elegans studies of rescue of longevity and fat storage in daf-2(e1370); atg-18(gk378) double mutants by tissue-specific expression of atg-18(+) revealed a major role of atg-18 in food-sensing chemosensory neurons, potentially reflecting a vesicle trafficking role in neurosecretion (possibly of neuropeptides) (Jia et al., 2019; Minnerly et al., 2017). Whether such a role is unique to atg-18 or shared by other atg genes remains unclear.

Third, recent evidence suggests that ATG-18 activates HLH-30/TFEB (Helix Loop Helix/Transcription Factor EB). C. elegans HLH-30 is a master transcriptional regulator of autophagy and lysosomal biogenesis, that becomes nuclear localized in daf-2(e1370) and glp-1(e2141) mutants, and whose inhibition reduces longevity in both contexts (Lapierre et al., 2013; Lin et al., 2018; Wong et al., 2023). Notably, among both genes up-regulated upon over-expression of atg-18, and genes down-regulated in an atg-18(gk378) mutant, ones with HLH-30 elements in their promoters are over-represented (Schmauck-Medina et al., 2026). Also, heat shock-induced nuclear localization of HLH-30 is suppressed by atg-18 RNAi, while longevity resulting from atg-18 over-expression is suppressed by hlh-30 RNAi.

Inhibition of HLH-30 by atg-18 RNAi could potentially explain its unusually strong life-shortening effects. If correct, this could reflect greater suppression of autophagy, but also, in principle, other effects of HLH-30 inhibition. HLH-30 affects a variety of traits, including adult reproductive diapause (Gerisch et al., 2020), proteostasis (Shalash et al., 2025), sex differences in immunity (Sohn et al., 2025), and resistance to various insults, including infectious pathogens (El-Houjeiri et al., 2019; Visvikis et al., 2014; Wani et al., 2021), starvation (Harvald et al., 2017; O’Rourke and Ruvkun, 2013; Settembre et al., 2013), and oxidative and heat stress (Lin et al., 2018).

While in at least some of these cases autophagy likely contributes to hlh-30-mediated effects, given that this transcription factor also promotes expression of many genes unrelated to autophagy (Chen et al., 2017; Lin et al., 2018; Shalash et al., 2025; Sohn et al., 2025; Visvikis et al., 2014), other functions may also be involved. For example, in response to infection HLH-30 can activate expression of signaling (including IIS), autophagy-related and immunity-related genes, and both of latter can contribute to infection resistance (Chen et al., 2017; Sohn et al., 2025; Visvikis et al., 2014). In conclusion, if the more severe effects of atg-18 RNAi on lifespan reported here are attributable to reduced HLH-30 activity, this could well reflect greater reduction of autophagy. However, it remains possible that other functions controlled by HLH-30 also play a role, e.g. relating to immunity (Chen et al., 2017; Visvikis et al., 2014) and protein quality control (Shalash et al., 2025). Also notable is that HLH-30 regulates gene expression combinatorially with DAF-16 (Lin et al., 2018), mutation of which fully suppresses daf-2(rf) longevity (Kenyon et al., 1993).

Finally, recent findings suggest that dependence of glp-1 Age on atg-18 is in part unrelated to its role in autophagy, but instead to an alternative pathway involving PCK-2 phosphoenolpyruvate carboxykinase (Shioda et al., 2026). Taken together, emerging findings argue that one may not safely conclude that effects of atg-18 on lifespan are attributable solely to its role in autophagy.

Autophagy in biomass repurposing during programmatic aging

Autophagic processes play a major role in tissue atrophy and degeneration related to biomass repurposing in semelparous organisms (that reproduce once and then die), particularly plants (Avila-Ospina et al., 2014; Gems et al., 2021). Previous studies support the hypothesis that intestinal biomass is repurposed for synthesis of yolk that, subsequent to egg laying, is vented to support larval development (Kern et al., 2021), and that autophagy facilitates this biomass conversion (Ezcurra et al., 2018; Sornda et al., 2019). If late-life mortality is promoted by intestinal atrophy, then preventing it should extend lifespan. Consistent with this, intestinal atrophy and yolk production are suppressed in daf-2 mutants (Depina et al., 2011; Ezcurra et al., 2018), and blocking yolk production both retards intestinal atrophy and modestly increases lifespan (Ezcurra et al., 2018; Murphy et al., 2003; Sornda et al., 2019), in part by increasing resistance to late-life E. coli infection (Wang et al., 2026). This could imply that inhibiting autophagy, by retarding intestinal atrophy, could extend lifespan under some conditions, and several earlier studies observed life extension after atg RNAi (Ezcurra et al., 2018; Hashimoto et al., 2009; Wilhelm et al., 2017).

The results of the present study are in line with the view that intestinal atrophy, though a salient feature of C. elegans senescence, by itself contributes only weakly to late-life mortality. Consistent with this, inhibition of daf-2 using auxin-induced degradation of DAF-2 protein can strongly increasing lifespan even when initiated only at advanced ages, after intestinal atrophy has occurred, and without any detectable reversal of major senescent pathologies (Molière et al., 2024; Venz et al., 2021). One possibility is that atg RNAi has antagonistic effects on late-life mortality: modestly reducing it due to suppression of intestinal atrophy (as seen when vitellogenin synthesis is inhibited) (Ezcurra et al., 2018; Murphy et al., 2003; Sornda et al., 2019), but also increasing it, due to disruption of essential cellular functions.

The gut-to-yolk biomass repurposing hypothesis drew particularly on the observation that blocking yolk protein (vitellogenin) synthesis or autophagy reduced intestinal atrophy and pseudocoelomic yolk pool size (Ezcurra et al., 2018; Sornda et al., 2019); given that yolk pools contain yolk protein and lipid (Ezcurra et al., 2018; Garigan et al., 2002; Herndon et al., 2002; Palikaras et al., 2017), such putative biomass repurposing could increase production of yolk protein, yolk lipid or both. Here we tested whether atg RNAi reduces yolk protein levels, and found that it did not (Figure 6). If the repurposing hypothesis is correct, then this would support the view that promotion by autophagy of repurposing of intestinal biomass into yolk involves yolk lipid rather than yolk protein, or overall yolk secretion. In principle, this could involve mass export of stored lipid, either by means of lipophagy, or perhaps secretory autophagy (Ponpuak et al., 2015). As an aside, we note that there is evidence that vitellogenin levels can affect autophagy (Seah et al., 2016).

Concluding remarks

This study reassesses the claim that daf-2 Age is autophagy dependent. While the results do not entirely prove or disprove this claim, they clearly show that evidence of dependence varies greatly with context, and suggest that autophagy only contributes substantially to daf-2 Age given a weak reduction in IIS.

One limitation in establishing the role of autophagy in daf-2 mutant longevity is the difficulty of reliably measuring its level of activity. While several assays using fluorescent reporters to identify putative autophagosomes suggest possible increases in autophagy levels (Chang et al., 2017; Hansen et al., 2008; Meléndez et al., 2003), daf-2 mutants also show marked reductions in both protein synthesis rate (Depuydt et al., 2013) and turnover rate (Dhondt et al., 2016; Visscher et al., 2016), including reduced ATG-18 turnover rate (Visscher et al., 2016). This is more consistent with a reduction rather than an increase in autophagy, and in line with observed suppression of intestinal atrophy in daf-2 mutants (Ezcurra et al., 2018).

One issue raised by condition dependence of a given effect is that it creates a risk of condition selection bias, where choice of experimental conditions can bias the outcome of an experiment. Where condition dependence has been identified, condition selection bias may be avoided by selection of multiple conditions to test a given hypothesis. In the present case, this could entail comparing effects on daf-2(e1368) and daf-2(e1370) at 20°C, ideally without FUDR use, and RNAi tests with multiple atg genes, but not atg-18, at least until the reason for its idiosyncratic effects are understood. Another factor to take into account is the apparent tendency of results of C. elegans lifespan assays to vary over time (Lucanic et al., 2017). General questions relating to how to deal with condition dependency have been discussed previously (Munafò and Davey Smith, 2018; Voelkl et al., 2020).

A second, wider issue is that condition dependence risks generating a situation where different research articles make conflicting claims, supporting different views held by different groups of researchers. For science to progress effectively it is necessary for research communities to resolve discrepancies between published findings, though the work required can be tedious, a task akin to washing the dishes in a communal household. It requires identifying and flagging findings that are either condition dependent or, seemingly, unreproducible and, where possible, distinguishing the two. For C. elegans at least, such a tidying process is highly feasible, as is use of methodologies devised to reduce such discrepancies (Driscoll et al., 2025; Lucanic et al., 2017; Petrascheck and Miller, 2017), which is a particular virtue of this model organism. More widely, condition dependency and conditional selection bias risk diminishing the reliability of research findings in many scientific disciplines.

Materials and methods

Culture methods and strains

C. elegans maintenance was performed using standard protocols (Brenner, 1974). Unless otherwise stated, all strains were on nematode growth media (NGM, containing Bacto Peptone) with plates seeded with E. coli OP50 to provide a food source. An N2 hermaphrodite stock recently obtained from the Caenorhabditis Genetics Center was used as wild type (N2H) (Zhao et al., 2019). Genotypes of most mutants used are as described in Wormbase (www.wormbase.org). Strains used included CB4027 glp-1(e2141), GA633 daf-2(m577); wuIs177 [Pftn-1::gfp lin-15(+)], GA643 daf-16(mgDf50); daf-2(m577); wuIs177 [Pftn-1::gfp lin-15(+)]], GA1930 daf-2(e1370), GA1945 daf-2(m41), GA1960 daf-2(e1368), and JK574 fog-2(q71).

Constitutive dauer larva formation assay

daf-2(m41) animals were maintained on RNAi feeding strains for at least two generations prior to analysis. For the dauer larva formation assay, performed at 22°C, 10 L4-stage hermaphrodites were transferred onto RNAi plates and allowed to lay eggs for approximately 6 h, after which adults were removed. Dauers were scored at 166 h.

Epifluorescence microscopy

Nematodes were anaesthetized with 10 μl 2 mM levamisole on 2% agar pads prior to imaging. For imaging, we used either a Zeiss Axioskop 2 plus microscope with Hamamatsu ORCA-ER digital camera C4742-95 and Volocity 6.3 software (Macintosh version) for image acquisition; or an ApoTome.2 Zeiss microscope with a Hamamatsu digital camera C13440 ORCA-Flash4.0 V3 and Zen software for image acquisition.

RNA-mediated interference (RNAi)

RNAi by feeding was performed using RNAi plasmids transformed into E. coli OP50(xu363) as previously described (Xiao et al., 2015). Bacterial transformants were selected on LB agar plates containing 10 µg/ml tetracycline and 25 µg/ml carbenicillin. Inserts of all RNAi feeding clones were confirmed by sequencing. Origins of plasmids in RNAi feeding strains: atg-13, bec-1: Ahringer library (Kamath et al., 2003); atg-2, atg-4.1, atg-9: Vidal library (Rual et al., 2004). As an atg-18 RNAi feeding clone was not available in local RNAi libraries, a custom atg-18 RNAi plasmid was generated by cloning the target sequence into the L4440 vector using primers gcctccacttcctgttgaag and gagactcttttcgtcggca.

For RNAi induction, bacteria were grown for 16 h at 37°C with shaking (200 rpm) in 5 ml LB supplemented with 25 µg/ml carbenicillin. Overnight cultures were diluted 1:100 into fresh LB containing 25 µg/ml carbenicillin and grown for a further 4 h or until reaching an OD600 of ∼0.4. Expression of double-stranded RNA was induced by addition of 1 mM IPTG, and cultures were incubated for an additional 1 h. Following induction, cultures were allowed to cool to room temperature, concentrated five-fold by centrifugation, and seeded onto NGM plates supplemented with 25 µg/ml carbenicillin and 1 mM IPTG. Seeded plates were allowed to dry at room temperature before use.

To minimize IPTG degradation, RNAi plates were stored in foil-lined boxes. Typically, unseeded plates were stored at 4°C for up to one month, while seeded plates were used within two weeks. RNAi treatment was initiated from the L4 larval stage unless otherwise stated.

Survival analysis

Nematodes were maintained at a density of 25-30 per plate, and transferred daily during the egg laying period, and every 6-7 days thereafter. The L4 stage was defined as day 0. Mortality was scored every 1-2 days, with worms scored as alive if they showed any movement, either spontaneously or in response to gentle touch with a worm pick.

RNA extraction, cDNA synthesis and RT-qPCR

Approximately 100–200 day 4 adult animals per treatment group were lysed in 250 µl of TRIzol™ reagent by vortexing for 10 min at 4°C, followed by a 10 min incubation on ice; this cycle was repeated three times. RNA was purified using the RNeasy Mini Kit (QIAGEN, Cat. No. 74104) according to the manufacturer’s instructions, including on-column DNase digestion using the RNase-Free DNase Set (QIAGEN, Cat. No. 79256). cDNA was synthesized from 100 ng of total RNA using the SuperScript™ First-Strand Synthesis System (Thermo Fisher Scientific, Cat. No. 10684803).

PCR primers were designed using Primer-BLAST with the following parameters: maximum primer length of 20 nucleotides, PCR product size of 70–200 bp, and exon–exon spanning where possible. Primer efficiency and R² values were determined from standard curves to ensure suitability for quantitative analysis and comparable amplification efficiencies between primer pairs. Primers were used at a final concentration of 2 µM.

Quantitative PCR was performed using Fast SYBR™ Green chemistry on a QuantStudio™ 6 Flex Real-Time PCR System (Applied Biosystems) with 384-well plates and a total reaction volume of 10 µl. Cycling conditions were 95°C for 20 s, followed by 40 cycles of 95°C for 1 s and 65°C for 20 s, with fluorescence data acquisition during the annealing/extension step. Melt curve analysis was performed following amplification to confirm product specificity.

A full list of primers used in this study is provided in Table S11. mRNA levels were normalized to mRNA levels from two housekeeping genes, cdc-42 and pmp-3 (Hoogewijs et al., 2008).

Electrophoresis of C. elegans yolk protein

N2 and fog-2(q71) worms were synchronized by performing an egg lay and allowing nematodes to grow at 20°C until they reached the L4 stage. L4 worms were then transferred to fresh seeded RNAi plates and maintained under standard culture conditions at 20°C. Five worms were harvested at day 1, day 4 and day 7 into microcentrifuge Eppendorf tubes containing 10 μl M9 buffer. Before running the gel, 10 μl of 2x Laemmli sample buffer was added. Samples were incubated at 70°C and vortexed periodically for 15 min. Samples were then incubated at 95°C and vortexed periodically for 5 min, and were centrifuged at 6,000 rpm for 15 min.

10 μl of sample was loaded into wells of Invitrogen NuPAGE Bis-Tris protein gels. 5 μl of CozyHi pre-stained protein ladder was loaded at the left side of the gels. The running buffer used was 5% XT MOPS (Bio-Rad). The gels were then run at 150V for 2 hr. Gels were removed from cassette and placed in 100 ml of Coomassie staining solution overnight. Gels were then washed with distilled water three times, then placed into destaining solution and soaked for 40 min. Gels were then washed with distilled water and stored at 4°C until imaged. Gels were imaged and saved as 8 bit grayscale TIF files. Images of gels were analysed using Fiji software. Bands of interest (e.g. myosin, YP170, YP115, YP88) were selected manually based on their molecular weight, and intensity of the bands was measured and exported to Microsoft Excel for further analysis.

Data analysis and statistics

Data were plotted using ePrism 9.0 (GraphPad Software, Boston, MA, USA) or Jupyter Notebook using Python with the matplotlib, pandas and NumPy libraries. Statistical tests were performed on raw data using either Prism 9 or JMP Pro 15 (JMP Statistical Discovery LLC, Cary, NC, USA) unless otherwise stated. The specific tests and post hoc corrections performed are described in the figure legends. To detect alterations in lifespan, the log rank test was used. To compare the magnitude of changes in lifespan, Cox Proportional Hazard (CPH) analysis was used. RT-qPCR data was analyzed using the ΔΔCt method and t-tests. To compare yolk protein levels, a two-way ANOVA was used. For lifespan trials, no statistical methods were used to predetermine sample size. The experiments were not randomized. The investigators were not blinded to allocation during experiments and outcome assessment.